---
title: "Hypothesis Testing & Effect Size"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Did Community Policing Actually Improve Response Times?

The city of Westbrook implemented a new community policing program in January. The police chief claims the program has improved emergency response times. Before the program, the average response time was **8.2 minutes** with a standard deviation of **1.5 minutes**. You've collected data from 20 randomly selected emergency calls after the program started.

Your editor wants to know two things: Is the improvement real? And is it big enough to matter?

### Task 1: Load the tidyverse

```{r}
library(tidyverse)
```

### Task 2: Set up the data and the hypotheses (2 points)

Before we run any tests, we need to state our hypotheses. The null hypothesis is what we assume to be true until we have evidence otherwise — like "innocent until proven guilty." Fill in the REPLACE_ME values.

**Null Hypothesis (H₀):** The city of Westbrook's new community policing program has not improved emergency response times.

**Alternative Hypothesis (H₁):** The city of Westbrook's new community policing program has improved emergency response times.

```{r}
# Response times (in minutes) after the community policing program
response_times <- c(
  7.8, 8.1, 7.5, 7.9, 8.3,
  7.6, 7.4, 8.0, 7.7, 7.8,
  8.2, 7.3, 7.9, 7.5, 8.1,
  7.6, 7.8, 7.4, 7.7, 7.9
)

# Known baseline (before the program)
baseline_mean <- 8.2
baseline_sd <- 1.5

# Create a tibble
response_data <- tibble(
  call_number = 1:20,
  response_time = response_times
)

response_data
```

### Task 3: Calculate summary statistics (1 point)

Let's see what the new response times look like. Replace the REPLACE_ME values.

```{r}
summary_stats <- response_data |>
  summarize(
    sample_size = n(),
    sample_mean = mean(response_time),
    sample_sd = sd(response_time),
    standard_error = sample_sd / sqrt(20)
  )

summary_stats
```

**Answer this question:** The sample mean is lower than the baseline of 8.2 minutes. How much lower? Does that seem like a meaningful improvement to you, just based on the numbers? What would you want to know before deciding?

The sample mean is 0.425 minutes lower than the baseline. Just based on the numbers, this seems like a meaningful improvement to me. However, I would want to know the margin of error for the new response times and how it compares to the margin of error for the old response times.

```{r}

# the new standard deviation is relatively small

8.2 - 7.775
```

### Task 4: Perform a one-sample t-test (2 points)

Now we formally test the police chief's claim. Since we're testing whether response times *decreased*, we use a one-sided test (`alternative = "less"`). Replace the REPLACE_ME with the baseline mean we're comparing against.

```{r}
# Set our significance level

# alpha = ideal p-value (0.05)

# mu = previous standard (old baseline)

alpha <- 0.05

# Perform the t-test
t_test_result <- t.test(
  response_data$response_time,
  mu = 8.2,
  alternative = "less"
)

t_test_result
```

### Task 5: Interpret the t-test results (2 points)

Let's pull out the key numbers. Look at the p-value: is it less than our alpha of 0.05?

```{r}
# Extract key values
p_value <- t_test_result$p.value

results <- tibble(
  `Baseline mean` = paste(baseline_mean, "minutes"),
  `New mean` = paste(round(summary_stats$sample_mean, 2), "minutes"),
  `Difference` = paste(round(baseline_mean - summary_stats$sample_mean, 2), "minutes"),
  `p-value` = round(p_value, 6),
  `Significant?` = if_else(p_value < alpha, "Yes - reject null hypothesis", "No - fail to reject")
)

results
```

**Answer this question:** Based on the p-value, can we reject the null hypothesis? What does this tell the police chief — and what does it NOT tell him?

Based on the p-value, we can reject the null hypothesis. This tells the police chief that the new initiative hasn't NOT decreased response times, but it does not tell him how much they have improved, if at all.


### Task 6: Visualize the results

```{r}
#doesn't give us any real information

ggplot(response_data, aes(x = response_time)) +
  geom_histogram(binwidth = 0.2, fill = "steelblue", color = "white", alpha = 0.7) +
  geom_vline(xintercept = baseline_mean, color = "red", linetype = "dashed", size = 1.2) +
  geom_vline(xintercept = summary_stats$sample_mean, color = "darkgreen", size = 1.2) +
  annotate("text", x = baseline_mean + 0.05, y = 4,
           label = paste("Baseline:", baseline_mean),
           color = "red", hjust = 0, fontface = "bold") +
  annotate("text", x = summary_stats$sample_mean - 0.05, y = 3.5,
           label = paste("New Mean:", round(summary_stats$sample_mean, 2)),
           color = "darkgreen", hjust = 1, fontface = "bold") +
  labs(
    title = "Emergency Response Times: Before vs. After Community Policing",
    subtitle = paste0("p-value = ", round(p_value, 6)),
    x = "Response Time (minutes)",
    y = "Number of Calls"
  ) +
  theme_minimal()
```

### Task 7: Calculate the effect size (3 points)

The t-test told us the improvement is statistically significant — it's real, not due to chance. But is it *meaningful*? That's what effect size measures. We'll calculate **Cohen's d**, which tells us how large the difference is in standard deviation units.

The formula is: d = (baseline mean - sample mean) / baseline standard deviation

Replace the REPLACE_ME values.

```{r}

# Cohen's d is always between 0 and 1 (the larger it is, the larger the effect size is)

effect_size <- (baseline_mean - summary_stats$sample_mean) / baseline_sd

effect_size_interpretation <- case_when(
  abs(effect_size) < 0.2 ~ "Negligible",
  abs(effect_size) < 0.5 ~ "Small",
  abs(effect_size) < 0.8 ~ "Medium",
  TRUE ~ "Large"
)

tibble(
  `Cohen's d` = round(effect_size, 3),
  `Interpretation` = effect_size_interpretation,
  `Difference in minutes` = round(baseline_mean - summary_stats$sample_mean, 2),
  `Difference in seconds` = round((baseline_mean - summary_stats$sample_mean) * 60, 0)
)
```

**Answer this question:** The result is statistically significant, but the effect size is small. In practical terms, the improvement is about 25 seconds per call. The police chief is claiming success. Would you write a story saying the program works? How would you frame the findings honestly — acknowledging both that the improvement is real AND that it's small?

I would write a story saying that the program works, but I would note that it has had a small effect. I would say something, either in the headline or lede, along the lines of, "New policing program slightly improves response times."


### Task 8: The journalist's framework (5 points)

The videos you watched describe two questions you need to ask about any statistical result:

1. **Is it statistically significant?** (Is the result unlikely to be due to chance?)
2. **Is it meaningfully significant?** (Is the effect large enough to matter?)

Apply this framework to the community policing data:

**Question 1:** Write 1-2 sentences answering whether the result is statistically significant and how you know.

The result is statistically significant. I know this because the sample's p-value is much lower than 0.05, meaning there is an extremely low chance that the change in police response times was due to chance.


**Question 2:** Write 1-2 sentences answering whether the result is meaningfully significant. Consider: the program cost the city $2 million to implement. Is a 25-second improvement in response times worth that investment? What context would you need to answer that question?

The result is not meaningfully significant because I do not believe that a 25-second improvement in police response times is worth a $2 million investment. I would need to know the current costs associated with policing in the city, as well as how much delayed response times cost the city. Additionally, I would need to know if there have been any cases where a short delay in police response times led to serious injury or death.


**Question 3:** Write a headline and a 2-3 sentence lede for a story about these findings. Your text should accurately reflect both the statistical significance and the effect size — don't let the police chief's framing drive the story if the data tells a more nuanced one.


New Westbrook policing program slightly improves emergency response times (headline)

The new Westbrook community policing program has slightly improved emergency response times. The average response time has improved by 25 seconds, according to a sample of 20 calls taken after January, when the program began. (lede)


When you are finished, save your work, switch to GitHub Desktop, then add, commit and push your changes to GitHub and submit the URL of the notebook in ELMS.
