---
title: "JOUR405: Statistics for Journalists"
subtitle: "Midterm Exam"
name: Josephine Andoh
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# JOUR405: Statistics for Journalists
## Midterm Exam

For this exam, you'll analyze several datasets using R and the statistical concepts we've covered in class. Load the tidyverse before beginning, then complete each task. Write your code in the provided blocks and answer the questions in complete sentences.

## Load the tidyverse

```{r}
library(tidyverse)
```

---

## Part 1: Restaurant Health Inspections (25 points)

The first dataset contains restaurant health inspection scores from Montgomery County. Load the data from: `https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/montco_inspections.csv`

The dataset includes the name of the establishment, the number of points for critical and non-critical areas, the total points, maximum points possible, the compliance score, and the grade.

### Task 1: Read the data (2 points)

```{r}
health <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/montco_inspections.csv")
```

### Task 2: Calculate the mean, standard deviation, minimum, and maximum of compliance scores. Save the results in a new dataframe called `compliance_stats`. (2 points for each)

```{r}
compliance_stats <- health |>
  summarize (
    mean_compliance = mean(compliance_score),
    sd_compliance = sd(compliance_score),
    min_compliance = min(compliance_score),
    max_compliance = max(compliance_score)
  )
```

### Task 3: Create a histogram of the compliance scores with a vertical line showing the mean. Include a descriptive title and label the x and y axes using `labs()`. (10 points)

```{r}
health |> 
  ggplot() + 
  geom_histogram(aes(x = compliance_score), binwidth = 10) +
  geom_vline(aes(xintercept = mean(compliance_score)), color = "red", linetype = "dashed", linewidth = 1) +
  theme_minimal() +
  labs(title = "Majority of Montgomery County Restaurants are Majorly Compliant",
       x = "Compliance Score",
       y = "Number of Restaurants") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### Task 4: Write 2–3 sentences interpreting what the standard deviation and histogram tell us about the distribution of compliance scores. What is newsworthy, and what would your next reporting step be? (5 points)

**Answer:**

The standard deviation and histogram tell us that the compliance scores have a high negative skew. They also tell us that most of the restaurants have compliance scores ranging from 90 to 100%, meaning that they have high compliance. The newsworthy aspect of this datset are the small number of restaurants with compliance scores between 65 and 75%, since they are much lower than the average. My next reporting step would be to identify those restaurants and see if any health incidents contributed to those relatively low scores.

---

## Part 2: Daily Transit Ridership — Z-Scores and Sampling (30 points)

Your editor wants you to identify unusual days in WMATA's daily ridership data that might be worth investigating as story leads. The dataset contains daily bus and rail totals. Load the data from: `https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/wmata_daily.csv`

### Task 5: Load the WMATA ridership data. (2 points)

```{r}
wmata <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/wmata_daily.csv")
```

### Task 6: Using `summarize()`, calculate the mean and standard deviation for both bus and rail ridership. Save the results in a dataframe called `wmata_stats`. (5 points)

```{r}
wmata_stats <- wmata |>
  summarize (
    rail_mean = mean(rail),
    rail_sd = sd(rail), 
    bus_mean = mean(bus),
    bus_sd = sd(bus)
  )
```

### Task 7: Using `mutate()`, add two new columns to `wmata_daily`: `rail_zscore` (the z-score for rail ridership) and `bus_zscore` (the z-score for bus ridership). Save the result as `wmata_zscores`. (8 points)

```{r}
wmata_zscores <- wmata |>
  mutate(
    rail_zscore = (rail - mean(rail)) / sd(rail),
    bus_zscore = (bus - mean(bus)) / sd(bus)
)
```

### Task 8: Filter `wmata_zscores` to show only days where the absolute value of `rail_zscore` is greater than 2 (unusually high or low rail ridership). Arrange by the absolute value of the z-score, with the most extreme first. How many unusual days are there? (8 points)

```{r}
unusual_rail_zscores <- wmata_zscores |>
  filter(abs(rail_zscore) > 2) |>
  arrange(desc(abs(rail_zscore)))
```

There are six unusual days.


### Task 9: Using `set.seed()`, take a random sample of 91 days from `wmata_daily` (roughly 25% of the data). Calculate the mean and standard deviation for bus and rail ridership in the sample. How do the sample statistics compare to the population statistics you calculated in Task 6? Write 1–2 sentences comparing the values and explaining what this suggests about random sampling. (4 points)

```{r}
set.seed(42)

sample_91 <- wmata |> sample_n(91)

sample_91_results <- sample_91 |>
  summarize(
    mean_rail = mean(rail),
    sd_rail = sd(rail),
    mean_bus = mean(bus),
    sd_bus = sd(bus)
  )
```

The mean and standard deviation for both the rail and bus ridership are higher in the sample than they are in the population, but except for the standard deviation of bus ridership, the samples are relatively close to the population. This suggests that random sampling can be an accurate way to analyze large datasets, but it will rarely be completely accurate.


### Task 10: Write 2–3 sentences: What do the unusual days in your z-score analysis suggest as story leads? (3 points)

The unusual days in my z-score analysis suggest that rail ridership is unusually low in early January and on winter holidays (I am assuming that November 28, 2024, was Thanksgiving Day). The low ridership in January could be a story lede because despite the workforce returning around those dates, rail ridership is relatively low, which could be caused by bad weather or rail track issues due to cold weather. The low ridership during the holidays, on the other hand, is less newsworthy.


---

## Part 3: Maryland Car Theft Rates (25 points)

Your editor has assigned you a story about car thefts in Maryland and wants you to find out which counties have the highest rates. The dataset contains car theft counts for Maryland counties in 2023 along with population figures. Load the data from: `https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_car_thefts.csv`

### Task 11: Load the car thefts data. (2 points)

```{r}
md_car_thefts <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_car_thefts.csv")
```

### Task 12: Using `mutate()`, add a column called `theft_rate` that calculates the car theft rate per 10,000 residents for each county. Save the result back to `md_car_thefts`. In 1–2 sentences, explain why a per-capita rate is more useful than raw counts for this comparison. (8 points)

```{r}
md_car_thefts <- md_car_thefts |>
  mutate (theft_rate_2023 = thefts_2023/population *10000)
```

A per-capita rate is more useful than raw counts for this comparison because it ensures that we are comparing different population sizes on equal bases, meaning that no areas will appear to have larger or smaller car theft rates than they actually do.


### Task 13: Calculate the median car theft rate and the total number of car thefts statewide. Show the counties that have rates above the median. (7 points)

```{r}
md_car_thefts <- md_car_thefts |>
  mutate (
    median_rate = median(theft_rate_2023))

md_car_thefts |> summarize (
  total = sum(thefts_2023)
)

md_car_thefts |>
  filter (county > median_rate)

md_car_thefts |>
  arrange(desc(theft_rate_2023))

md_car_thefts |>
  filter (theft_rate_2023 > 9.580638)
```

### Task 14: Write 2–3 sentences describing what these calculations reveal about the distribution of car thefts in Maryland. What is the lede of a story based on your findings? (8 points)

These calculations reveal that 50% of large Maryland districts had car theft rates in 2023 that were above the median. However, they also show that Baltimore City, Prince George's County and St. Mary's County had very high car theft rates and would have skewed our analysis if we had used the mean to analyze instead.

Based on my findings, a possible lede of a story is:

Baltimore City and St. Mary's County accounted for the majority of car thefts among Maryland's jurisdictions in 2023, according to state crime data.

---

## Part 4: Data Analysis Scenario (20 points)

A city parks department releases data showing that its 12 community swimming pools average 187 visitors per day during the summer season, with a standard deviation of 42 visitors. Your editor suspects that a handful of pools are dramatically underperforming and wants a story. You also learn that the department's figures are based on a random sample of 60 days per pool rather than a complete count.

Write 3–4 sentences for each of the following questions:

### Question A: Given that the mean is 187 and standard deviation is 42, which statistical measures would you prioritize using and why? (7 points)

Given that the mean is 187 and standard deviation is 42, I would prioritize making a histogram of the data. A histogram would help me visualize the data and calculate z-scores, which would help me see how closely the data follows the empirical rule. Using the empirical rule would help me detect and analyze any anomalies, such as if a pool had an extremely high or low number of visitors compared to the average. Also, since this is a relatively large dataset, this would greatly speed up analysis.


### Question B: How would you use z-scores to identify which pools are genuinely underperforming, and why is that more useful than simply ranking pools by raw attendance? (7 points)

I would use z-scores to identify which pools are genuinely underperforming by calculating them and then filtering based on which pools have a z-score of two or greater. I would also use z-scores to calculate pnorms, which would help me analyze which pools are supposed to have greater or lesser attendance based on probability. This is more useful than simply ranking pools by raw attendance because it objectively compares the pools to each other, rather than by numbers that appear large out of context.


### Question C: The department collected its attendance figures using a random sample of 60 days per pool rather than tracking every day. What questions would you ask about the sample before relying on these figures for a story? What concerns would you raise in your story? Use statistical terminology where appropriate. (6 points)

I would ask: Did you run any simulations based on that sample size? Why did you choose a sample of 60? What did the distribution of the samples look like? Did some of the sample days include when pools were closed?

I would be concerned because samples do not always accurately reflect reality if they are too small, and pool season goes on for roughly 3.5 to 4 months (or around 120 days). A sample of 60 days may not be large enough to incorporate ebbs and flows in data, such as bad weather in certain neighborhoods.

---

### Submission Instructions
- Save your work frequently
- Make sure all code blocks run without errors
- Provide clear explanations for your choices

Good luck!
