---
title: "Identifying Schools with Biggest Differences in Math Test Scores, Part 2"
author: "Josephine Andoh"
---

```{r}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
options(scipen = 999)
```

### Introduction

In the class activity, you tested a school board member's claim using data from Montgomery County, Pa. — and found a model that didn't hold up. Now you'll run the same analysis using data from Montgomery County, Md. Your job is to evaluate whether the Maryland data tells a different story, and what that means for how you'd report on the claim.

### Loading the Data

```{r}
schools <- read_csv("https://raw.githubusercontent.com/dwillis/jour405_files/refs/heads/main/montgomery_md.csv")
```

### Task 1: Examine the Data First (2 points)

Before you plot anything, look at the data — you know why by now.

```{r}
summary(schools$`FARMS Pct`)
schools |> arrange(desc(`FARMS Pct`)) |> head(10)
```

Are there any data quality issues like the ones you found in the Pennsylvania data? How does the distribution of FARMS percentages compare between the two counties? Note anything that concerns you or gives you confidence in the data.


Based on initial observation, there are few (if any) quality issues with the Montgomery County, Maryland, data. There are no schools with a 0 or 100% share of FARMS-eligible students, and the data has a realistic range (4 to 96%). 

Additionally, I grew up and currently live in Montgomery County, and the shares of FARMS-eligible students seem realistic. For example, Broad Acres Elementary School is in a lower-income area of the county, so it makes sense that it has a high share of FARMS-eligible students. On the other hand, Bethesda Elementary School is in a higher-income area of the county, so it makes sense that it has a low share of FARMS-eligible students.


### Task 2: Create a Scatterplot and Compare (5 points)

Replace the REPLACE_ME value with the correct variable name to create a scatterplot of FARMS percentage vs. math scores for Montgomery County, Md.

```{r}
schools |>
  ggplot(aes(x = `FARMS Pct`, y = `% Met or Exceeded Math Expectations`)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(x = "FARMS Percentage", y = "% Met or Exceeded Math Expectations",
       title = "Higher share of FARMS-eligible students at Montgomery County elementary schools may predict worse student math performance, data shows")
```

Describe what you see. Then compare this scatterplot specifically to the Pennsylvania one: what differences do you notice in the direction, strength, and spread of the relationship? The two counties share a name and are both suburban — what might explain why they look so different?


The scatterplot appears to show that the more FARMS-eligible students an elementary school has, the lower the share of students who meet or exceed math expectations.

This scatterplot and the Pennsylvania county's scatterplot are rather different. The Maryland county's scatterplot shows a strong negative relationship, with most data points close to the line of best fit. On the other hand, the Pennsylvania county's scatterplot shows a very weak positive relationship, with most data points widely dispersed.

Something that might explain why the two scatterplots look so different is how FARMS eligibility was calculated. I suspect the Pennsylvania county's calculations were incorrect in some way, because it is very unrealistic that there are eight elementary schools where 100% of students are FARMS-eligible. On the other hand, I believe the Maryland county's calculations were mostly (if not completely) accurate because they are realistic (based on personal knowledge and the data's range).


### Task 3: Linear Regression Model (5 points)

Replace the REPLACE_ME value with the correct variable name to run a linear regression model predicting math scores from FARMS percentage.

```{r}
model <- lm(`% Met or Exceeded Math Expectations` ~ `FARMS Pct`, data = schools)
summary(model)
```

Report the three key statistics: the p-value, the Adjusted R-squared, and the coefficient for FARMS Pct. Then interpret them together — not separately. What does the combination of these three numbers tell you about the relationship between poverty and math scores in Montgomery County, Md.? Is this model worth using? How does it compare to the Pennsylvania model?


The p-value for this data is a number less than 0.00000000000000022. The adjusted R-squared value for this data is 0.6734. The coefficient for the FARMS percentage is -0.59511. 

The combination of these three numbers tells me that there is a somewhat strong, statistically significant relationship between poverty and math scores in elementary schools in Montgomery County, Maryland. More specifically, these three numbers tell me that an increase in the number of students in poverty at a given elementary school can somewhat predict lower math scores at that school.

This model could be worth using, but other variables likely contribute to the apparent correlation between poverty and math scores. However, it is much more useful than the Pennsylvania model.


### Task 4: Identifying Schools with the Biggest Residuals (6 points)

Replace the REPLACE_ME values with the correct variable names to calculate predicted values, residuals, and absolute residuals for each school. Then arrange the schools in descending order of absolute residuals.

```{r}
schools <- schools |>
  mutate(
    predicted = predict(model),
    residual = `% Met or Exceeded Math Expectations` - predicted,
    abs_residual = abs(residual)
  ) |>
  arrange(desc(abs_residual))

head(schools |> select(1, `FARMS Pct`, `% Met or Exceeded Math Expectations`, predicted, residual), 10)
```

Identify the school with the largest positive residual (doing much better than the model predicts) and the school with the largest negative residual (doing much worse). For each school, find its zip code and look it up on [CensusReporter.org](https://censusreporter.org).

For each of the two schools, answer:
- What does the demographic and economic profile of the zip code look like?
- Does anything in that profile help explain why the school is over- or under-performing the model?
- What would your first reporting question be to the principal of that school?


The school with the largest positive residual is Fox Chapel Elementary School, located in zip code 20876. 29% of people living in that zip code are Black, 25% are Hispanic, 22% are white, 20% are Asian, 4% are multiracial and 1% are some other race. 51% of residents are female and 49% are male. The median age of residents is 36.8. Zip code 20876 has a per capita income of $47,977 and a median household income of $119,731. Additionally, 9.8% of residents live below the poverty line.

Something in the zip code's profile that might help explain its overperformance is the education level of its residents. 88.6% have graduated from high school or obtained a higher level of education, and 52% have received a bachelor's degree or higher. The more educated parents are, the more likely they are to help their children with their schoolwork, which can improve students' academic performance. My first reporting question to the principal of that school would be: How do math teachers at your school help underperforming students?


The school with the largest negative residual is Germantown Elementary School, located in zip code 20874. 27% of people living in that zip code are Black, 26% are white, 22% are Hispanic, 18% are Asian and 6% are biracial. 51% of residents are female and 49% are male. The median age of residents is 37.6. Zip code 20874 has a per capita income of $52,219 and a median household income of $





Population more than double 20876 - larger classrooms mean less individualized instruction



### Task 5: Evaluate the Model and Write the Pitch (7 points)

You've now run this analysis on two counties. Return to the school board member's claim: that elementary schools with fewer students on free and reduced-price meals do better on standardized math tests.

Write a short story pitch of 4-6 sentences that addresses all of the following: Does the Maryland data support or contradict the claim? How does it compare to what you found in Pennsylvania? What would you ask the school board member if you called them for comment? And what's the most interesting specific finding from your residuals analysis: which school would you want to visit first, and why?


ANSWER HERE

