---
title: "HW3 WH Salaries"
---

In this assignment, we will be working with a dataset containing information about the salaries of White House staffers as of July 2025.

### Task 1: Load the Tidyverse

You will need to load the tidyverse library

```{r}
library(tidyverse)
```

Then you'll load the data from the web. The URL for the data is: https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/white_house_salaries_2025.csv. You will create a dataframe called `wh_salaries` that contains the data from this URL.

### Task 2: Load the Data and Describe It

```{r}
wh_salaries <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/white_house_salaries_2025.csv")
```

### Task 3: Describe the Data

You should examine that dataframe carefully by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the categories of incidents. From a reporting standpoint, what stands out to you about this data? Are there any salary values that will cause problems for analysis?


The two categories of employment statuses are employee and detailee. The categories of position titles are managers, assistants, directors, writers, speechwriters, personal aides, counsels, coordinators, policy advisors, stenographers, lead advance representatives, associates, shift leads, supervisors, analysts, editors, operators and czars. 

From a reporting standpoint, what stands out to me about this data is how high Jacalynne Klopp's salary is, since their job title is not very specific. Salaries of $0 could cause problems for analysis because they will be excluded from both mean and median calculations. Those values are also confusing since it is unlikely that employees at any institution, especially not the White House, would work for free.


### Task 4: Calculate the Mean and Median of the Salaries

Write code to calculate the mean and median of the salaries in the dataset using the `summarize()` function. You don't need to make a new dataframe, just display the results.

```{r}
wh_salaries|>
  summarize(
    mean_salary = mean(`SALARY`)
    )

wh_salaries|>
  summarize(
    median_salary = median(`SALARY`)
  )
```

### Task 5: Add in the minimum and maximum salaries

Take the code you just wrote and add in the minimum and maximum salaries (the functions for that are `min()` and `max()`). Your results should show four columns: mean_salary, median_salary, min_salary, max_salary.

```{r}
wh_salaries|>
  summarize(
    mean_salary = mean(`SALARY`)
    )

wh_salaries|>
  summarize(
    median_salary = median(`SALARY`)
  )
    
wh_salaries|>
  summarize(
    max_salary = max(`SALARY`)
  )

wh_salaries|>
  summarize(
    min_salary = min(`SALARY`)
  )
```

### Task 6: Describe the Results

Write a couple of sentences describing the results of your analysis. What do the mean and median salaries tell you about the data? What do the minimum and maximum salaries tell you? Are there any outliers in the data?

The mean and median salaries tell me that the data are positively skewed. The fact that the mean is roughly $8,000 higher than the median shows that there are more salaries above $109,117.40 than below. The minimum salary tells me that a handful of White House employees work in unpaid positions. The maximum salary tells me that one White House detailee makes more money than all other detailees and employees, including Marco Rubio and Karoline Leavitt. Consequently, the highest salary could be an outlier in the data.


### Task 7: Story Idea

Based on the data, what is your best story idea? What is the next step you would take to develop that story idea?

Based on the data, my best story idea is that as of July 2025, the highest salary at the Trump White House is $225,700 per year. The next step I would take to develop this story idea is to research Jacalynne Klopp and figure out why their salary is so high. Ideally, I would stumble upon an interesting piece of information and use it to write a story that gives people more insight into the Trump White House.
