---
title: "HW4 MD CityCrime"
---

In this assignment, we will be working with a dataset of total crimes reported by Maryland cities for two consecutive years. You'll load the data, calculate the violent and property crime rates for each city for both years, and then calculate the percentage change in the crime rates between the two years. You'll also write a short story based on your findings.

### Task 1: Load the Tidyverse and Janitor

You will need to load the tidyverse library and the janitor library

```{r}
library(tidyverse)
library(janitor)
```

Then you'll load the data from the web. The URL for the data is: https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_cities_crime_2019_2020.csv. You will create a dataframe called `md_cities` that contains the data from this URL.

### Task 2: Load the Data

```{r}
md_cities <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_cities_crime_2019_2020.csv")
```

### Task 3: Calculating Rates

Calculate violent and property crime rates for each city for both years using `mutate()`. You'll create four new columns: `violent_rate_2019`, `property_rate_2019`, `violent_rate_2020`, and `property_rate_2020`. You'll need to figure out the proper unit rate, given that there are varying populations. There's no perfect choice; choose the one that works best for you. Save the results to a new dataframe called `md_cities_rates`.

```{r}
md_cities_rates <- md_cities |> mutate(
  violent_rate_2019 = violent2019/pop2019 * 10000,
  property_rate_2019 = property2019/pop2019 * 10000,
  violent_rate_2020 = violent2020/pop2020 * 10000,
  property_rate_2020 = property2020/pop2020 * 10000
)
```

### Task 4: Describe the Rates

Examine the new dataframe, `md_cities_rates`, by clicking on it in the Environment tab in RStudio and sorting by the rate columns. Write a couple of sentences about the results, comparing and contrasting the cities with the highest rates (both violent and property) in 2019 and 2020. Do you see any patterns? What does this tell you about crime in Maryland cities? What's the story here?


One pattern I see is that Ocean City had either the highest or second-highest violent and property crime rates of all major Maryland cities in 2019 and 2020. Another pattern I see is that Baltimore had either the highest or second-highest rate of violent crime of all major Maryland cities in 2019 and 2020. An additional pattern I see is that Oakland and Elkton had some of the highest property crime rates of all major Maryland cities in 2019 and 2020. 

This tells me that the higher property and crime rates in Maryland in 2019 and 2020 were concentrated around a few major cities. The story here is that Ocean City had a consistently high property and crime rate in 2019 and 2020, undermining the popular assumption that Baltimore is Maryland's "most dangerous" city.


### Task 5: Calculate Percentage Change

Let's add two percentage change columns to your dataframe: one for violent crime rates and one for property crime rates. You'll calculate the percentage change between 2019 and 2020 for each city. Save the results to a new dataframe called `md_cities_change`.

```{r}
md_cities_change <- md_cities_rates |> mutate(
  pct_change_violent = ((violent_rate_2020 - violent_rate_2019)/violent_rate_2019 * 100),
  pct_change_property = ((property_rate_2020 - property_rate_2019)/property_rate_2019 * 100)
  )
```

### Task 7: Describe Percentage Changes

Look at the `md_cities_change` dataframe by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the percentage change results. What's more interesting here, the patterns or the outliers? What's the story in the data?


Chevy Chase Sector 5, Laytonsville and Friendsville had the largest increase in violent crime rates from 2019 to 2020. Cecilton, Queenstown and New Windsor had the largest decrease in violent crime rates from 2019 to 2020. Keedysville, Burkittsville and Sharpsburg had the largest increase in property crime rates from 2019 to 2020. Friendship Heights, Brookeville and Church Hill had the largest decrease in property crime rates from 2019 to 2020. 

To me, the patterns in this dataset are more interesting than the outliers. From 2019 to 2020, violent crime rates increased in 58 major Maryland cities, decreased in 47 and stayed the same in 31. During the same period, property crime rates decreased in 78 cities, increased in 44 and stayed the same in 15. To me, the story in the data is that violent crime rates in a plurality of major Maryland cities increased from 2019 to 2020, and that property crime rates in a majority of those cities decreased during the same period.


### Task 8: Tell the Best Story

Write up a lede and nutgraf of a story with the _best_ story idea you've found in this assignment. The lede should be a single sentence that summarizes the most important information in the data. The nutgraf should be a paragraph that provides more context and explains why the data is important.


Ocean City had the highest property and violent crime rates of all major Maryland cities in 2020, according to state crime data.

Approximately 240 violent crimes and 1,213 property crimes were committed per 10,000 residents, dwarfing rates in cities such as Baltimore and Annapolis. Ocean City also had the highest property crime rate and second-highest violent crime rate of all major Maryland cities in 2019. Between the two years, the city's violent crime rates increased by approximately 83%, and its property crime rates increased by roughly 17%.


### Task 9: Save Your Work and Put it on GitHub

Save this notebook, then commit it to your local repository. Push the changes to your GitHub repository and submit the URL for this file in ELMS.
