---
title: "HW6_Spotify_Songs"
---

In this assignment, you will be working with a dataset of songs played on Spotify in 2020, and you'll be writing code to explore the range of danceability scores in the data, using standard deviation and histograms. You'll also write up your findings.

### Task 1: Load the tidyverse

```{r}
library(tidyverse)
```

### Task 2: Load the Spotify Songs dataset from https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/spotify_2020.csv

```{r}
songs <- read.csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/spotify_2020.csv")
```

### Task 3: Calculate the mean and standard deviation of the danceability scores

Write code to calculate the mean and standard deviation of the danceability scores in the Spotify dataset. Display those.

```{r}
songs |> summarize(mean = mean(danceability), sd = sd(danceability))
```

```{r}
songs |> summarize(mean(danceability) - sd(danceability))
songs |> summarize(mean(danceability) + sd(danceability))
```


### Task 4: Create a histogram of the danceability scores

Write code to create a histogram of the danceability scores in the Spotify dataset, based on the example we did in class.

```{r}
songs |> 
  ggplot() +
  geom_histogram(aes(x = danceability), binwidth = 0.15)
```

### Task 5: Add a line to the histogram for the mean danceability score

Take the same code and add a vertical line to the histogram for the mean danceability score, based on the example in class.

```{r}
songs |> 
  ggplot() +
  geom_histogram(aes(x = danceability), binwidth = 0.15) +
  geom_vline(aes(xintercept = mean(danceability)), color = "red", linetype = "dashed", size = 1)
```

### Task 6: Add two lines to the histogram to show one standard deviation

Take the previous code and add two lines to the histogram for the 1st standard deviation below and above the mean score.

```{r}
songs |> 
  ggplot() +
  geom_histogram(aes(x = danceability), binwidth = 0.15) +
  geom_vline(aes(xintercept = mean(danceability)), color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = mean(danceability) - sd(danceability)), color = "blue", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = mean(danceability) + sd(danceability)), color = "blue", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = mean(danceability) - sd(danceability) - sd(danceability)), color = "blue", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = mean(danceability) + sd(danceability) + sd(danceability)), color = "blue", linetype = "dashed", size = 1)
```

### Task 7: Describe our results and pitch a story

Write a few sentences about the distribution of danceability scores in the Spotify dataset. Is it a normal distribution, or is it skewed in one direction? What does the mean and one standard distribution tell us about the songs in the dataset? Draft a short story pitch from this data.


The danceability scores in the Spotify dataset are negatively skewed. Though the average danceability score was roughly 67.3%, and most of the scores were within one standard deviation from the mean, a handful of the scores were more than two standard deviations lower than it. This tells us that even though the average Spotify song in 2020 had a relatively high danceability score, some songs were not very danceable.

Based on this data, this would be my story pitch: 

The average Spotify song had a danceability score of approximately 67.3% in 2020. However, some songs had very low danceability scores. How does Spotify determine danceability? Do measures such as energy, instrumentalness and liveness factor into the danceability score, and if so, how? How did artists and music producers react to their music's danceability scores?


### Task 8: Save Your Work and Put it on GitHub

Save this notebook, then commit it to your local repository. Push the changes to your GitHub repository and submit the URL for this file in ELMS.
