---
title: "final_project"
author: Josephine Andoh
output: html_document
---


## I loaded the tidyverse, knitr and lubridate so I can analyze my datasets.

```{r}
library(tidyverse)

knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(knitr) 

library(lubridate)
```


## I loaded crime data from Prince George's County from 2017 to 2023 and 2023 to 2026, then combined those datasets to facilitate my analysis. I then loaded crime data from Montgomery County from 2017 to 2026.

```{r}
pgcrime17_23 <- read_csv("PGCRIME17-23.csv")

pgcrime23_26 <- read_csv("PGCRIME23-26.csv")

mococrime <- read_csv("MOCOCRIME.zip")

pgcrime <- bind_rows(pgcrime17_23, pgcrime23_26)
```


## I sorted all the crimes into seven categories (assault, burglary, auto theft, theft from auto, robbery, sex offenses and homicide) to facilitate my analysis.

```{r}
pgcrime <- pgcrime |> 
  mutate(category = case_when(
    str_detect(`Clearance Code Inc Type`, "ASSAULT") ~ "assault",
    str_detect(`Clearance Code Inc Type`, "ASSAULT, WEAPON") ~ "assault",
    str_detect(`Clearance Code Inc Type`, "ASSAULT, SHOOTING") ~ "assault",
    str_detect(`Clearance Code Inc Type`, "B & E, RESIDENTIAL") ~ "burglary",
    str_detect(`Clearance Code Inc Type`, "B & E, RESIDENTIAL (VACANT)") ~ "burglary",
    str_detect(`Clearance Code Inc Type`, "B & E, COMMERCIAL") ~ "burglary",
    str_detect(`Clearance Code Inc Type`, "B & E, OTHER") ~ "burglary",
    str_detect(`Clearance Code Inc Type`, "AUTO, STOLEN") ~ "auto theft",
    str_detect(`Clearance Code Inc Type`, "AUTO, STOLEN & RECOVERED") ~ "auto theft",
    str_detect(`Clearance Code Inc Type`, "ROBBERY, OTHER") ~ "robbery",
    str_detect(`Clearance Code Inc Type`, "ROBBERY, COMMERCIAL") ~ "robbery",
    str_detect(`Clearance Code Inc Type`, "ROBBERY, RESIDENTIAL") ~ "robbery",
    str_detect(`Clearance Code Inc Type`, "ROBBERY, VEHICLE") ~ "robbery",
    str_detect(`Clearance Code Inc Type`, "ROBBERY, VEHICLE") ~ "robbery",
    str_detect(`Clearance Code Inc Type`, "THEFT") ~ "theft",
    str_detect(`Clearance Code Inc Type`, "THEFT FROM AUTO") ~ "theft from auto",
    str_detect(`Clearance Code Inc Type`, "SEX OFFENSE") ~ "sex offense",
    str_detect(`Clearance Code Inc Type`, "HOMICIDE") ~ "homicide",
    .default = NA
  )
)


mococrime <- mococrime |>
  mutate(category = case_when(
    str_detect(`Crime Name3`, "ASSAULT - 2ND DEGREE") ~ "assault",
    str_detect(`Crime Name3`, "ASSAULT - INTIMIDATION (INCLUDES STALKING)") ~ "assault",
    str_detect(`Crime Name3`, "ASSAULT - AGGRAVATED - PUB OFF-OTHER WEAPON") ~ "assault",
    str_detect(`Crime Name3`, "ASSAULT - AGGRAVATED - POL OFF-OTHER WEAPON") ~ "assault",
    str_detect(`Crime Name3`, "ASSAULT - AGGRAVATED - FAMILY-GUN") ~ "assault",
    str_detect(`Crime Name3`, "ASSAULT - SIMPLE") ~ "assault",
    str_detect(`Crime Name3`, "ASSAULT - AGGRAVATED - GUN") ~ "assault",
    str_detect(`Crime Name3`, "ASSAULT - AGGRAVATED - NON-FAMILY-OTHER WEAPON") ~ "assault",
    str_detect(`Crime Name3`, "ASSAULT - AGGRAVATED - FAMILY-STRONG-ARM") ~ "assault",
    str_detect(`Crime Name3`, "ASSAULT - AGGRAVATED - NON-FAMILY-GUN") ~ "assault",
    str_detect(`Crime Name3`, "ASSAULT - AGGRAVATED - NON-FAMILY-STRONG-ARM") ~ "assault",
    str_detect(`Crime Name3`, "ATTEMPTED ASSAULT AGGRAVATED POL OFF KNIFE") ~ "assault",
    str_detect(`Crime Name3`, "ASSAULT - AGGRAVATED - POL OFF-STRONG-ARM") ~ "assault",
    str_detect(`Crime Name3`, "ASSAULT - AGGRAVATED - POL OFF-GUN") ~ "assault",
    str_detect(`Crime Name3`, "ASSAULT - AGGRAVATED - PUB OFF-STRONG-ARM") ~ "assault",
    str_detect(`Crime Name3`, "ASSAULT - AGGRAVATED - PUB OFF-GUN") ~ "assault",
    
    .default = NA
  )
)

mococrime <- mococrime |>
  mutate(category = case_when(
    str_detect(`Crime Name3`, "SEX OFFENSE - AGAINST CHILD-FONDLING") ~ "sex offense",
    str_detect(`Crime Name3`, "SEX ASSAULT - FONDLING") ~ "sex offense",
    str_detect(`Crime Name3`, "SEX OFFENSE - PEEPING TOM") ~ "sex offense",
    str_detect(`Crime Name3`, "SEX OFFENSE - INDECENT EXPOSURE") ~ "sex offense",
    str_detect(`Crime Name3`, "COMM SEX OFF - KEEPING HOUSE ILL FAME") ~ "sex offense",
    str_detect(`Crime Name3`, "SEX OFFENSE - INCEST WITH ADULT") ~ "sex offense",
    str_detect(`Crime Name3`, "SEX ASSAULT - SODOMY-GIRL-GUN") ~ "sex offense",
    str_detect(`Crime Name3`, "SEX ASSAULT - SODOMY-GIRL-WEAPON") ~ "sex offense",
    str_detect(`Crime Name3`, "COMM SEX OFF - PROCURE FOR PROSTITUTE (PIMPING)") ~ "sex offense",
    str_detect(`Crime Name3`, "COMM SEX OFF - PROCURE PROSTITUTE - MINOR") ~ "sex offense",
    str_detect(`Crime Name3`, "SEX ASSAULT - SODOMY-WOMAN-GUN") ~ "sex offense",
    str_detect(`Crime Name3`, "COMM SEX OFF - PROSTITUTION") ~ "sex offense",
    str_detect(`Crime Name3`, "HOMOSEXUAL ACT WITH BOY") ~ "sex offense",
    str_detect(`Crime Name3`, "SEX OFFENSE - INCEST WITH MINOR") ~ "sex offense",
    str_detect(`Crime Name3`, "SEX ASSAULT - SODOMY-MAN-WEAPON") ~ "sex offense",
    str_detect(`Crime Name3`, "COMM SEX OFF - HOMOSEXUAL PROSTITUTION") ~ "sex offense",
    str_detect(`Crime Name3`, "SEX ASSAULT - SODOMY-WOMAN-WEAPON") ~ "sex offense",
    str_detect(`Crime Name3`, "SEX ASSAULT - SODOMY-MAN-GUN") ~ "sex offense",
    str_detect(`Crime Name3`, "COMM SEX OFF - TRANSPORT FEMALE INTERSTATE FOR IMM") ~ "sex offense",
    .default = NA
  )
)

mococrime <- mococrime |>
  mutate(category = case_when(
    str_detect(`Crime Name3`, "ROBBERY - STREET-GUN") ~ "robbery",
    str_detect(`Crime Name3`, "ROBBERY - RESIDENTIAL-STRONG-ARM") ~ "robbery",
    str_detect(`Crime Name3`, "ROBBERY - STREET-STRONG-ARM") ~ "robbery",
    str_detect(`Crime Name3`, "ROBBERY - STREET-OTHER WEAPON") ~ "robbery",
    str_detect(`Crime Name3`, "ROBBERY - BUSINESS-GUN") ~ "robbery",
    str_detect(`Crime Name3`, "ROBBERY - FORCIBLE PURSE SNATCHING") ~ "robbery",
    str_detect(`Crime Name3`, "ROBBERY (DESCRIBE OFFENSE)") ~ "robbery",
    str_detect(`Crime Name3`, "ROBBERY - RESIDENTIAL-GUN") ~ "robbery",
    str_detect(`Crime Name3`, "ROBBERY - BUSINESS-OTHER WEAPON") ~ "robbery",
    str_detect(`Crime Name3`, "ROBBERY - CARJACKING - ARMED") ~ "robbery",
    str_detect(`Crime Name3`, "ROBBERY - RESIDENTIAL-OTHER WEAPON") ~ "robbery",
    str_detect(`Crime Name3`, "ROBBERY - BANKING-TYPE INST") ~ "robbery",
    .default = NA
  )
)

mococrime <- mococrime |>
  mutate(category = case_when(
    str_detect(`Crime Name3`, "BURGLARY - SAFE-VAULT") ~ "burglary",
    str_detect(`Crime Name3`, "BURGLARY - BANKING-TYPE INST") ~ "burglary",
    str_detect(`Crime Name3`, "BURGLARY (DESCRIBE OFFENSE)") ~ "burglary",
    str_detect(`Crime Name3`, "BURGLARY - NO FORCED ENTRY-NONRESIDENTIAL") ~ "burglary",
    str_detect(`Crime Name3`, "BURGLARY - FORCED ENTRY-RESIDENTIAL") ~ "burglary",
    str_detect(`Crime Name3`, "BURGLARY - FORCED ENTRY-NONRESIDENTIAL") ~ "burglary",
    str_detect(`Crime Name3`, "BURGLARY - NO FORCED ENTRY-RESIDENTIAL") ~ "burglary",
    .default = NA
  )
)

mococrime <- mococrime |>
  mutate(category = case_when(
    str_detect(`Crime Name3`, "STOLEN VEHICLE (DESCRIBE OFFENSE)") ~ "auto theft",
    str_detect(`Crime Name3`, "AUTO THEFT - VEHICLE THEFT") ~ "auto theft",
    str_detect(`Crime Name3`, "AUTO THEFT - THEFT VEHICLE BY BAILEE") ~ "auto theft",
    str_detect(`Crime Name3`, "AUTO THEFT - STRIP STOLEN VEHICLE") ~ "auto theft",
    str_detect(`Crime Name3`, "AUTO THEFT - THEFT AND STRIP VEHICLE") ~ "auto theft",
    str_detect(`Crime Name3`, "AUTO THEFT - THEFT AND SALE VEHICLE") ~ "auto theft",
    str_detect(`Crime Name3`, "AUTO THEFT - THEFT AND USE VEHICLE OTHER CRIME") ~ "auto theft",
    .default = NA
  )
)

mococrime <- mococrime |>
  mutate(category = case_when(
    str_detect(`Crime Name3`, "LARCENY - FROM AUTO") ~ "theft from auto",
    str_detect(`Crime Name3`, "LARCENY - AUTO PARTS") ~ "theft from auto",
    .default = NA
  )
)

mococrime <- mococrime |>
  mutate(category = case_when(
    str_detect(`Crime Name3`, "HOMICIDE - WILLFUL KILL-GUN") ~ "homicide",
    str_detect(`Crime Name3`, "HOMICIDE (DESCRIBE OFFENSE)") ~ "homicide",
    str_detect(`Crime Name3`, "HOMICIDE - WILLFUL KILL") ~ "homicide",
    str_detect(`Crime Name3`, "HOMICIDE - WILLFUL KILL-FAMILY") ~ "homicide",
    str_detect(`Crime Name3`, "HOMICIDE - JUSTIFIABLE") ~ "homicide",
    str_detect(`Crime Name3`, "HOMICIDE - WILLFUL KILL-FAMILY-GUN") ~ "homicide",
    str_detect(`Crime Name3`, "HOMICIDE - WILLFUL KILL-NON-FAMILY-GUN") ~ "homicide",
    str_detect(`Crime Name3`, "HOMICIDE - WILLFUL KILL-NON-FAMILY") ~ "homicide",
    str_detect(`Crime Name3`, "HOMICIDE - NEGLIGENT MANSLAUGHTER") ~ "homicide",
    str_detect(`Crime Name3`, "HOMICIDE - BY NON-POLICE - JUSTIFIABLE") ~ "homicide",
    str_detect(`Crime Name3`, "HOMICIDE - JOHN OR JANE DOE - NO WARRANT") ~ "homicide",
    .default = NA
  )
)

mococrime <- mococrime |>
  mutate(category = case_when(
    str_detect(`Crime Name3`, "LARCENY (DESCRIBE OFFENSE)") ~ "theft",
    str_detect(`Crime Name3`, "LARCENY - POCKET PICKING") ~ "theft",
    str_detect(`Crime Name3`, "LARCENY - FROM INTERSTATE SHIPMENT") ~ "theft",
    str_detect(`Crime Name3`, "LARCENY - THEFT OF US GOVERNMENT PROPERTY") ~ "theft",
    str_detect(`Crime Name3`, "LARCENY - FROM COIN MACHINE") ~ "theft",
    str_detect(`Crime Name3`, "LARCENY - FROM SHIPMENT") ~ "theft",
    str_detect(`Crime Name3`, "LARCENY - FROM BANKING-TYPE INST") ~ "theft",
    str_detect(`Crime Name3`, "LARCENY - FROM MALLS") ~ "theft",
    str_detect(`Crime Name3`, "LARCENY - FROM YARDS") ~ "theft",
    str_detect(`Crime Name3`, "LARCENY - PURSE SNATCHING - NO FORCE") ~ "theft",
    str_detect(`Crime Name3`, "LARCENY - POCKET PICKING") ~ "theft",
    str_detect(`Crime Name3`, "LARCENY - POSTAL") ~ "theft",
    str_detect(`Crime Name3`, "LARCENY - FROM BLDG") ~ "theft",
    str_detect(`Crime Name3`, "LARCENY - SHOPLIFTING") ~ "theft",
    .default = NA
  )
)
```


## I filtered my data (to the best of my ability) on a monthly basis so that I was not comparing too many individual data points.

```{r}
pgcrime <- pgcrime |>
  mutate(Date = mdy(Date)) |> 
  mutate(
    month = floor_date(Date, "month")
  )
```

```{r}
mococrime <- mococrime |> 
  mutate(Date = date(mdy_hm(Start_Date_Time)))
```


## I filtered my data (to the best of my ability) from April 1, 2017, to April 1, 2026, so that I could compare crime rates over the same time period.

```{r}
new_pgcrime <- pgcrime |> filter(Date < '2026-04-02')

new_pgcrime <- pgcrime |> filter(Date > '2017-03-31')
```

```{r}
new_mococrime <- mococrime |> filter(Start_Date_Time > '3/31/2017 23:59')

new_mococrime <- mococrime |> filter(Start_Date_Time < '4/1/2026 0:00')
```


## Step 5: Calculating crime rates

```{r}

```


New questions:

Do I have to upload population data so I can calculate crime rates?




Hypotheses:

1. Montgomery County has higher rates of burglaries, automobile thefts and thefts from automobiles than Prince George's County
2. Prince George's County has higher rates of assault, sex offenses and overall theft than Montgomery County
3. Prince George's County has higher overall rates of these six crime categories than Montgomery County


The Tests I Could Run:

An analysis I think would be useful for my project is calculating correlations among the different crime categories I am analyzing. It would be interesting to see if burglary and theft rates, for example, were correlated. However, it would be even more interesting if two crime categories that, at first glance, could be correlated (automobile thefts and thefts from automobiles, for example) actually had no relationship.

Another analysis I think would be useful for my project is calculating the correlation between assault and overall crime rates for Prince George's and Montgomery counties. You mentioned once in class that assault rates can reflect overall crime rates, and it would be interesting to see if that held in those counties over a long time period.

A third analysis I think would be useful for my project is calculating the z-scores of the number of crimes committed per month and analyzing any unusual z-scores. Seeing which county has more months with unusually high z-scores could make my project more interesting and would prompt me to research those months to determine the reasons for the increases in crime.

A fourth analysis I think would be useful for my project is calculating whether any decreases or increases in crime were statistically significant. If I found that they were, I could also calculate their effect size. This would give me more information than z-scores would, and could also prompt me to do more research.
