---
title: "final_project"
author: Josephine Andoh
output: html_document
---


For my final project, I analyzed per capita crime rates in Prince George's and Montgomery counties between 2017 and April 2026. My hypotheses were:

1. Montgomery County had higher rates of burglaries, automobile thefts and thefts from automobiles than Prince George's County during this time period.

2. Prince George's County had higher rates of assault, sex offenses and overall theft than Montgomery County during this time period.

3. Prince George's County had higher overall rates of these six crime categories than Montgomery County during this time period.



## Step 1: I loaded the tidyverse, knitr, lubridate, dplyr and scales so I could analyze and graph my datasets.

```{r}
library(tidyverse)

knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(knitr) 

library(lubridate)

library(dplyr)

library(scales)
```



## Step 2: I loaded crime data from Prince George's County from 2017 to 2023 and 2023 to 2026, then combined those datasets to facilitate my analysis. I then loaded crime data from Montgomery County from 2016 to 2026.

```{r}
pgcrime17_23 <- read_csv("PGCRIME17-23.csv")

pgcrime23_26 <- read_csv("PGCRIME23-26.csv")

mococrime <- read_csv("MOCOCRIME.zip")

pgcrime <- bind_rows(pgcrime17_23, pgcrime23_26)
```



## Step 3: I sorted the relevant crimes into eight categories (assault, burglary, automobile theft, theft from automobiles, robbery, sex offenses, theft and homicide) based on general Maryland law to facilitate my analysis. 

```{r}

# Categorizing Prince George's County's data. 
pgcrime <- pgcrime |> 
  mutate(category = case_when(
    str_detect(`Clearance Code Inc Type`, "ASSAULT") ~ "assault",
    str_detect(`Clearance Code Inc Type`, "ASSAULT, WEAPON") ~ "assault",
    str_detect(`Clearance Code Inc Type`, "ASSAULT, SHOOTING") ~ "assault",
    str_detect(`Clearance Code Inc Type`, "B & E, RESIDENTIAL") ~ "burglary",
    str_detect(`Clearance Code Inc Type`, "B & E, RESIDENTIAL (VACANT)") ~ "burglary",
    str_detect(`Clearance Code Inc Type`, "B & E, COMMERCIAL") ~ "burglary",
    str_detect(`Clearance Code Inc Type`, "B & E, OTHER") ~ "burglary",
    str_detect(`Clearance Code Inc Type`, "AUTO, STOLEN") ~ "auto theft",
    str_detect(`Clearance Code Inc Type`, "AUTO, STOLEN & RECOVERED") ~ "auto theft",
    str_detect(`Clearance Code Inc Type`, "ROBBERY, OTHER") ~ "robbery",
    str_detect(`Clearance Code Inc Type`, "ROBBERY, COMMERCIAL") ~ "robbery",
    str_detect(`Clearance Code Inc Type`, "ROBBERY, RESIDENTIAL") ~ "robbery",
    str_detect(`Clearance Code Inc Type`, "ROBBERY, VEHICLE") ~ "robbery",
    str_detect(`Clearance Code Inc Type`, "ROBBERY, VEHICLE") ~ "robbery",
    str_detect(`Clearance Code Inc Type`, "THEFT FROM AUTO") ~ "theft from auto",
    str_detect(`Clearance Code Inc Type`, "THEFT") ~ "theft",
    str_detect(`Clearance Code Inc Type`, "SEX OFFENSE") ~ "sex offense",
    str_detect(`Clearance Code Inc Type`, "HOMICIDE") ~ "homicide",
    .default = NA
  )
)


# Categorizing Montgomery County's assault data.
mococrime <- mococrime |>
  mutate(category = case_when(
    str_detect(`Crime Name3`, "ASSAULT - 2ND DEGREE") ~ "assault",
    str_detect(`Crime Name3`, "ASSAULT - INTIMIDATION (INCLUDES STALKING)") ~ "assault",
    str_detect(`Crime Name3`, "ASSAULT - AGGRAVATED - PUB OFF-OTHER WEAPON") ~ "assault",
    str_detect(`Crime Name3`, "ASSAULT - AGGRAVATED - POL OFF-OTHER WEAPON") ~ "assault",
    str_detect(`Crime Name3`, "ASSAULT - AGGRAVATED - FAMILY-GUN") ~ "assault",
    str_detect(`Crime Name3`, "ASSAULT - SIMPLE") ~ "assault",
    str_detect(`Crime Name3`, "ASSAULT - AGGRAVATED - GUN") ~ "assault",
    str_detect(`Crime Name3`, "ASSAULT - AGGRAVATED - NON-FAMILY-OTHER WEAPON") ~ "assault",
    str_detect(`Crime Name3`, "ASSAULT - AGGRAVATED - FAMILY-STRONG-ARM") ~ "assault",
    str_detect(`Crime Name3`, "ASSAULT - AGGRAVATED - NON-FAMILY-GUN") ~ "assault",
    str_detect(`Crime Name3`, "ASSAULT - AGGRAVATED - NON-FAMILY-STRONG-ARM") ~ "assault",
    str_detect(`Crime Name3`, "ATTEMPTED ASSAULT AGGRAVATED POL OFF KNIFE") ~ "assault",
    str_detect(`Crime Name3`, "ASSAULT - AGGRAVATED - POL OFF-STRONG-ARM") ~ "assault",
    str_detect(`Crime Name3`, "ASSAULT - AGGRAVATED - POL OFF-GUN") ~ "assault",
    str_detect(`Crime Name3`, "ASSAULT - AGGRAVATED - PUB OFF-STRONG-ARM") ~ "assault",
    str_detect(`Crime Name3`, "ASSAULT - AGGRAVATED - PUB OFF-GUN") ~ "assault",
    
    .default = NA
  )
)

# Categorizing Montgomery County's sex offense data.
mococrime <- mococrime |>
  mutate(category = case_when(
    str_detect(`Crime Name3`, "SEX OFFENSE - AGAINST CHILD-FONDLING") ~ "sex offense",
    str_detect(`Crime Name3`, "SEX ASSAULT - FONDLING") ~ "sex offense",
    str_detect(`Crime Name3`, "SEX OFFENSE - PEEPING TOM") ~ "sex offense",
    str_detect(`Crime Name3`, "SEX OFFENSE - INDECENT EXPOSURE") ~ "sex offense",
    str_detect(`Crime Name3`, "COMM SEX OFF - KEEPING HOUSE ILL FAME") ~ "sex offense",
    str_detect(`Crime Name3`, "SEX OFFENSE - INCEST WITH ADULT") ~ "sex offense",
    str_detect(`Crime Name3`, "SEX ASSAULT - SODOMY-GIRL-GUN") ~ "sex offense",
    str_detect(`Crime Name3`, "SEX ASSAULT - SODOMY-GIRL-WEAPON") ~ "sex offense",
    str_detect(`Crime Name3`, "COMM SEX OFF - PROCURE FOR PROSTITUTE (PIMPING)") ~ "sex offense",
    str_detect(`Crime Name3`, "COMM SEX OFF - PROCURE PROSTITUTE - MINOR") ~ "sex offense",
    str_detect(`Crime Name3`, "SEX ASSAULT - SODOMY-WOMAN-GUN") ~ "sex offense",
    str_detect(`Crime Name3`, "COMM SEX OFF - PROSTITUTION") ~ "sex offense",
    str_detect(`Crime Name3`, "HOMOSEXUAL ACT WITH BOY") ~ "sex offense",
    str_detect(`Crime Name3`, "SEX OFFENSE - INCEST WITH MINOR") ~ "sex offense",
    str_detect(`Crime Name3`, "SEX ASSAULT - SODOMY-MAN-WEAPON") ~ "sex offense",
    str_detect(`Crime Name3`, "COMM SEX OFF - HOMOSEXUAL PROSTITUTION") ~ "sex offense",
    str_detect(`Crime Name3`, "SEX ASSAULT - SODOMY-WOMAN-WEAPON") ~ "sex offense",
    str_detect(`Crime Name3`, "SEX ASSAULT - SODOMY-MAN-GUN") ~ "sex offense",
    str_detect(`Crime Name3`, "COMM SEX OFF - TRANSPORT FEMALE INTERSTATE FOR IMM") ~ "sex offense",
    .default = category
  )
)

# Categorizing Montgomery County's robbery data.
mococrime <- mococrime |>
  mutate(category = case_when(
    str_detect(`Crime Name3`, "ROBBERY - STREET-GUN") ~ "robbery",
    str_detect(`Crime Name3`, "ROBBERY - RESIDENTIAL-STRONG-ARM") ~ "robbery",
    str_detect(`Crime Name3`, "ROBBERY - STREET-STRONG-ARM") ~ "robbery",
    str_detect(`Crime Name3`, "ROBBERY - STREET-OTHER WEAPON") ~ "robbery",
    str_detect(`Crime Name3`, "ROBBERY - BUSINESS-GUN") ~ "robbery",
    str_detect(`Crime Name3`, "ROBBERY - FORCIBLE PURSE SNATCHING") ~ "robbery",
    str_detect(`Crime Name3`, "ROBBERY (DESCRIBE OFFENSE)") ~ "robbery",
    str_detect(`Crime Name3`, "ROBBERY - RESIDENTIAL-GUN") ~ "robbery",
    str_detect(`Crime Name3`, "ROBBERY - BUSINESS-OTHER WEAPON") ~ "robbery",
    str_detect(`Crime Name3`, "ROBBERY - CARJACKING - ARMED") ~ "robbery",
    str_detect(`Crime Name3`, "ROBBERY - RESIDENTIAL-OTHER WEAPON") ~ "robbery",
    str_detect(`Crime Name3`, "ROBBERY - BANKING-TYPE INST") ~ "robbery",
    .default = category
  )
)

# Categorizing Montgomery County's burglary data.
mococrime <- mococrime |>
  mutate(category = case_when(
    str_detect(`Crime Name3`, "BURGLARY - SAFE-VAULT") ~ "burglary",
    str_detect(`Crime Name3`, "BURGLARY - BANKING-TYPE INST") ~ "burglary",
    str_detect(`Crime Name3`, "BURGLARY (DESCRIBE OFFENSE)") ~ "burglary",
    str_detect(`Crime Name3`, "BURGLARY - NO FORCED ENTRY-NONRESIDENTIAL") ~ "burglary",
    str_detect(`Crime Name3`, "BURGLARY - FORCED ENTRY-RESIDENTIAL") ~ "burglary",
    str_detect(`Crime Name3`, "BURGLARY - FORCED ENTRY-NONRESIDENTIAL") ~ "burglary",
    str_detect(`Crime Name3`, "BURGLARY - NO FORCED ENTRY-RESIDENTIAL") ~ "burglary",
    .default = category
  )
)

# Categorizing Montgomery County's automobile theft data.
mococrime <- mococrime |>
  mutate(category = case_when(
    str_detect(`Crime Name3`, "STOLEN VEHICLE (DESCRIBE OFFENSE)") ~ "auto theft",
    str_detect(`Crime Name3`, "AUTO THEFT - VEHICLE THEFT") ~ "auto theft",
    str_detect(`Crime Name3`, "AUTO THEFT - THEFT VEHICLE BY BAILEE") ~ "auto theft",
    str_detect(`Crime Name3`, "AUTO THEFT - STRIP STOLEN VEHICLE") ~ "auto theft",
    str_detect(`Crime Name3`, "AUTO THEFT - THEFT AND STRIP VEHICLE") ~ "auto theft",
    str_detect(`Crime Name3`, "AUTO THEFT - THEFT AND SALE VEHICLE") ~ "auto theft",
    str_detect(`Crime Name3`, "AUTO THEFT - THEFT AND USE VEHICLE OTHER CRIME") ~ "auto theft",
    .default = category
  )
)

# Categorizing Montgomery County's theft from automobile data.
mococrime <- mococrime |>
  mutate(category = case_when(
    str_detect(`Crime Name3`, "LARCENY - FROM AUTO") ~ "theft from auto",
    str_detect(`Crime Name3`, "LARCENY - AUTO PARTS") ~ "theft from auto",
    .default = category
  )
)

# Categorizing Montgomery County's homicide data.
mococrime <- mococrime |>
  mutate(category = case_when(
    str_detect(`Crime Name3`, "HOMICIDE - WILLFUL KILL-GUN") ~ "homicide",
    str_detect(`Crime Name3`, "HOMICIDE (DESCRIBE OFFENSE)") ~ "homicide",
    str_detect(`Crime Name3`, "HOMICIDE - WILLFUL KILL") ~ "homicide",
    str_detect(`Crime Name3`, "HOMICIDE - WILLFUL KILL-FAMILY") ~ "homicide",
    str_detect(`Crime Name3`, "HOMICIDE - JUSTIFIABLE") ~ "homicide",
    str_detect(`Crime Name3`, "HOMICIDE - WILLFUL KILL-FAMILY-GUN") ~ "homicide",
    str_detect(`Crime Name3`, "HOMICIDE - WILLFUL KILL-NON-FAMILY-GUN") ~ "homicide",
    str_detect(`Crime Name3`, "HOMICIDE - WILLFUL KILL-NON-FAMILY") ~ "homicide",
    str_detect(`Crime Name3`, "HOMICIDE - NEGLIGENT MANSLAUGHTER") ~ "homicide",
    str_detect(`Crime Name3`, "HOMICIDE - BY NON-POLICE - JUSTIFIABLE") ~ "homicide",
    str_detect(`Crime Name3`, "HOMICIDE - JOHN OR JANE DOE - NO WARRANT") ~ "homicide",
    .default = category
  )
)

# Categorizing Montgomery County's theft data.
mococrime <- mococrime |>
  mutate(category = case_when(
    str_detect(`Crime Name3`, "LARCENY (DESCRIBE OFFENSE)") ~ "theft",
    str_detect(`Crime Name3`, "LARCENY - POCKET PICKING") ~ "theft",
    str_detect(`Crime Name3`, "LARCENY - FROM INTERSTATE SHIPMENT") ~ "theft",
    str_detect(`Crime Name3`, "LARCENY - THEFT OF US GOVERNMENT PROPERTY") ~ "theft",
    str_detect(`Crime Name3`, "LARCENY - FROM COIN MACHINE") ~ "theft",
    str_detect(`Crime Name3`, "LARCENY - FROM SHIPMENT") ~ "theft",
    str_detect(`Crime Name3`, "LARCENY - FROM BANKING-TYPE INST") ~ "theft",
    str_detect(`Crime Name3`, "LARCENY - FROM MALLS") ~ "theft",
    str_detect(`Crime Name3`, "LARCENY - FROM YARDS") ~ "theft",
    str_detect(`Crime Name3`, "LARCENY - PURSE SNATCHING - NO FORCE") ~ "theft",
    str_detect(`Crime Name3`, "LARCENY - POCKET PICKING") ~ "theft",
    str_detect(`Crime Name3`, "LARCENY - POSTAL") ~ "theft",
    str_detect(`Crime Name3`, "LARCENY - FROM BLDG") ~ "theft",
    str_detect(`Crime Name3`, "LARCENY - SHOPLIFTING") ~ "theft",
    .default = category
  )
)
```

## Step 3 results: Through this sorting, I noticed that Montgomery County has much more specific crime categories than Prince George's County. I predicted that because of this, Montgomery County would appear to have higher crime rates per capita in certain categories, specifically assault.




## Step 4: I filtered my data to compare crime rates month-to-month. Any other time increment would have been too small or too large.

```{r}
pgcrime <- pgcrime |>
  mutate(Date = mdy(Date)) |> 
  mutate(
    month = floor_date(Date, "month")
  )
```

```{r}
mococrime <- mococrime |> 
  mutate(Date = date(mdy_hm(Start_Date_Time)))|>
      mutate(month = floor_date(Date, "month"))
```



## Step 5: I filtered my data from April 1, 2017, to March 31, 2026, so that I could compare crime rates over an exact nine-year time period.

```{r}
new_pgcrime <- pgcrime |> filter(Date < '2026-04-01', Date > '2017-03-31')
```

```{r}
new_mococrime <- mococrime |> filter(Date < '2026-04-01', Date > '2017-03-31')
```



## Step 6: I incorporated 2022 population data from both counties to calculate per-capita crime rates. I chose 2022 because it was close to the center of my data and would not reflect population changes caused by the COVID-19 pandemic. As I suspected, Montgomery County has a slightly larger population than Prince George's County.

## I retrieved the Montgomery County data from this link (https://montgomeryplanning.org/blog-design/2024/06/montgomery-countys-population-rebound-part-1-recovering-from-the-covid-19-years/) and the Prince George's County data from this link (https://www.princegeorgescountymd.gov/sites/default/files/media-document/County%20Profile_0.pdf).

```{r}
new_mococrime <- new_mococrime %>%
  mutate(population = 1053067)

new_pgcrime <- new_pgcrime %>%
  mutate(population = 946971)
```



## Step 7: I counted the number of specific crimes that occurred in both counties each month, then calculated per capita rates for each. I then combined the crime data from both counties into eight dataframes to facilitate my comparison. 

```{r}

# Counting the number of assaults in both counties and combining the dataframes.
pg_assault <- new_pgcrime |> 
  filter(category == "assault") |> 
  group_by(category, month, population) |> 
  summarise(pg_total = n()) |> 
  mutate(pg_assault_rate = pg_total/population * 100000)

moco_assault <- new_mococrime |> 
  filter(category == "assault") |> 
  group_by(category, month, population) |> 
  summarise(moco_total = n()) |> 
  mutate(moco_assault_rate = moco_total/population * 100000)

combined_assault <- pg_assault |> 
  inner_join(moco_assault, join_by(category, month))



# Counting the number of sex offenses in both counties and combining the dataframes.
pg_sexoffense <- new_pgcrime |> 
  filter(category == "sex offense") |> 
  group_by(category, month, population) |> 
  summarise(pg_total = n()) |> 
  mutate(pg_sexoffense_rate = pg_total/population * 100000)

moco_sexoffense <- new_mococrime |> 
  filter(category == "sex offense") |> 
  group_by(category, month, population) |> 
  summarise(moco_total = n()) |> 
  mutate(moco_sexoffense_rate = moco_total/population * 100000)

combined_sexoffense <- pg_sexoffense |> 
  inner_join(moco_sexoffense, join_by(category, month))



# Counting the number of robberies in both counties and combining the dataframes.
pg_robbery <- new_pgcrime |> 
  filter(category == "robbery") |> 
  group_by(category, month, population) |> 
  summarise(pg_total = n()) |> 
  mutate(pg_robbery_rate = pg_total/population * 100000)

moco_robbery <- new_mococrime |> 
  filter(category == "robbery") |> 
  group_by(category, month, population) |> 
  summarise(moco_total = n()) |> 
  mutate(moco_robbery_rate = moco_total/population * 100000)

combined_robbery <- pg_robbery |> inner_join(moco_robbery, join_by(category, month))



# Counting the number of burglaries in both counties and combining the dataframes.
pg_burglary <- new_pgcrime |> 
  filter(category == "burglary") |> 
  group_by(category, month, population) |> 
  summarise(pg_total = n()) |> 
  mutate(pg_burglary_rate = pg_total/population * 100000)

moco_burglary <- new_mococrime |> 
  filter(category == "burglary") |> 
  group_by(category, month, population) |> 
  summarise(moco_total = n()) |> 
  mutate(moco_burglary_rate = moco_total/population * 100000)

combined_burglary <- pg_burglary |> 
  inner_join(moco_burglary, join_by(category, month))



# Counting the number of automobile thefts in both counties and combining the dataframes.
pg_autotheft <- new_pgcrime |> 
  filter(category == "auto theft") |> 
  group_by(category, month, population) |> 
  summarise(pg_total = n()) |> 
  mutate(pg_autotheft_rate = pg_total/population * 100000)

moco_autotheft <- new_mococrime |> 
  filter(category == "auto theft") |> 
  group_by(category, month, population) |> 
  summarise(moco_total = n()) |> 
  mutate(moco_autotheft_rate = moco_total/population * 100000)

combined_autotheft <- pg_autotheft |> 
  inner_join(moco_autotheft, join_by(category, month))



# Counting the number of thefts from automobiles in both counties and combining the dataframes.
pg_theftfromauto <- new_pgcrime |> 
  filter(category == "theft from auto") |> 
  group_by(category, month, population) |> 
  summarise(pg_total = n()) |> 
  mutate(pg_theftfromauto_rate = pg_total/population * 100000)

moco_theftfromauto <- new_mococrime |> 
  filter(category == "theft from auto") |> 
  group_by(category, month, population) |> 
  summarise(moco_total = n()) |> 
  mutate(moco_theftfromauto_rate = moco_total/population * 100000)

combined_theftfromauto <- pg_theftfromauto |> 
  inner_join(moco_theftfromauto, join_by(category, month))



# Counting the number of homicides in both counties and combining the dataframes.
pg_homicide <- new_pgcrime |> 
  filter(category == "homicide") |> 
  group_by(category, month, population) |> 
  summarise(pg_total = n()) |> 
  mutate(pg_homicide_rate = pg_total/population * 100000)

moco_homicide <- new_mococrime |> 
  filter(category == "homicide") |> 
  group_by(category, month, population) |> 
  summarise(moco_total = n()) |> 
  mutate(moco_homicide_rate = moco_total/population * 100000)

combined_homicide <- pg_homicide |> 
  inner_join(moco_homicide, join_by(category, month))



# Counting the number of thefts in both counties and combining the dataframes.
pg_theft <- new_pgcrime |> 
  filter(category == "theft") |> 
  group_by(category, month, population) |> 
  summarise(pg_total = n()) |> 
  mutate(pg_theft_rate = pg_total/population * 100000)

moco_theft <- new_mococrime |> 
  filter(category == "theft") |> 
  group_by(category, month, population) |> 
  summarise(moco_total = n()) |> 
  mutate(moco_theft_rate = moco_total/population * 100000)

combined_theft <- pg_theft |> 
  inner_join(moco_theft, join_by(category, month))
```



## Step 8: I compared the rates for each crime between the two counties in graphs to more easily analyze the data. I noted my takeaways in the titles for each graph.

```{r}

# Assault graph.
combined_assault |> 
  ggplot() +
  geom_line((aes(x=month, y=pg_assault_rate)), color = "red", linewidth = 1) +
  geom_line(aes(x=month, y=moco_assault_rate), color = "blue", linewidth = 1) +
  labs(title = "Assault rates in Montgomery County, Maryland, consistently higher than in neighboring Prince George's County from early 2017 to early 2026, according to analysis",
       x = "Month",
       y = "Assault rate per 10,000 residents",
       caption = "Sources: Montgomery and Prince George's counties' crime archives")


# Automobile theft graph.
combined_autotheft |> 
  ggplot() +
  geom_line((aes(x=month, y=pg_autotheft_rate)), color = "red", linewidth = 1) +
  geom_line(aes(x=month, y=moco_autotheft_rate), color = "blue", linewidth = 1) +
  labs(title = "Auto theft rates in Prince George's County, Maryland, generally higher than in neighboring Montgomery County from early 2017 to early 2026, according to analysis",
       x = "Month",
       y = "Auto theft rate per 10,000 residents",
       caption = "Sources: Montgomery and Prince George's counties' crime archives")


# Burglary graph.
combined_burglary |> 
  ggplot() +
  geom_line((aes(x=month, y=pg_burglary_rate)), color = "red", linewidth = 1) +
  geom_line(aes(x=month, y=moco_burglary_rate), color = "blue", linewidth = 1) +
  labs(title = "Burglary rates in Prince George's and Montgomery counties in Maryland have decreased since 2017, according to analysis",
       x = "Month",
       y = "Burglary rate per 10,000 residents",
       caption = "Sources: Montgomery and Prince George's counties' crime archives")


# Homicide graph.
combined_homicide |> 
  ggplot() +
  geom_line((aes(x=month, y=pg_homicide_rate)), color = "red", linewidth = 1) +
  geom_line(aes(x=month, y=moco_homicide_rate), color = "blue", linewidth = 1) +
  labs(title = "Homicide rates in Prince George's County, Maryland, generally higher than in neighboring Montgomery County from early 2017 to early 2026, according to analysis",
       x = "Month",
       y = "Homicide rate per 10,000 residents",
       caption = "Sources: Montgomery and Prince George's counties' crime archives")


# Robbery graph.
combined_robbery |> 
  ggplot() +
  geom_line((aes(x=month, y=pg_robbery_rate)), color = "red", linewidth = 1) +
  geom_line(aes(x=month, y=moco_robbery_rate), color = "blue", linewidth = 1) +
  labs(title = "Robbery rates in Prince George's County, Maryland, generally higher than in neighboring Montgomery County from early 2017 to early 2026, according to analysis",
       x = "Month",
       y = "Robbery rate per 10,000 residents",
       caption = "Sources: Montgomery and Prince George's counties' crime archives")


# Sex offense graph.
combined_sexoffense |> 
  ggplot() +
  geom_line((aes(x=month, y=pg_sexoffense_rate)), color = "red", linewidth = 1) +
  geom_line(aes(x=month, y=moco_sexoffense_rate), color = "blue", linewidth = 1) +
  labs(title = "Sex offense rates in Montgomery County, Maryland, generally higher than in neighboring Prince George's County from early 2017 to early 2026, according to analysis",
       x = "Month",
       y = "Sex offense rate per 10,000 residents",
       caption = "Sources: Montgomery and Prince George's counties' crime archives")


# Theft graph.
combined_theft |> 
  ggplot() +
  geom_line((aes(x=month, y=pg_theft_rate)), color = "red", linewidth = 1) +
  geom_line(aes(x=month, y=moco_theft_rate), color = "blue", linewidth = 1) +
  labs(title = "Theft rates in Montgomery County, Maryland, generally higher than in neighboring Prince George's County from early 2017 to early 2026, according to analysis",
       x = "Month",
       y = "Theft rate per 10,000 residents",
       caption = "Sources: Montgomery and Prince George's counties' crime archives")


# Theft from automobile graph.
combined_theftfromauto |> 
  ggplot() +
  geom_line((aes(x=month, y=pg_theftfromauto_rate)), color = "red", linewidth = 1) +
  geom_line(aes(x=month, y=moco_theftfromauto_rate), color = "blue", linewidth = 1) +
  labs(title = "Theft from automobile rates in Prince George's and Montgomery counties in Maryland lower in early 2026 than in 2017",
       x = "Month",
       y = "Theft from automobile rate per 10,000 residents",
       caption = "Sources: Montgomery and Prince George's counties' crime archives")
```
## Step 8 results: 

## The result that surprised me the most from the graphic was Montgomery County's high assault rate compared to Prince George's County. It disproved my hypothesis and is an outlier compared to the other graphs, where the crime rates are closer together. I suspected that my categorization of assault may have contributed to this discrepancy. 

## It appears that automobile theft rates in both counties spiked in 2023. I have been hearing more about automobile thefts in recent years, so this piqued my interest and warrants further research.

## It appears that burglary rates in both counties have decreased overall since April 2017. This decrease warrants further research.

## It appears that homicide rates in Prince George's County spiked in late 2020 and in 2023. These spikes warrant further research.

## It appears that robbery rates in Prince George's County spiked in late 2023 and have steadily decreased since then. This change warrants further research.

## It appears that Prince George's County had a relatively high sex offense rate in 2017. However, since that year, both counties have sustained sex offense rates of less than five per 10,000 residents. However, to determine whether this number is high or low, I will have to compare it to the national average.

## It appears that theft rates in both counties had a spike in late 2024. This spike warrants further research.

## Theft from automobile rates in both counties appear to zigzag without a consistent pattern, which warrants further research.




## Step 9: I calculated the means, standard deviations and z-scores for each crime in their combined dataframes to determine whether the number of occurrences of each crime per month was anomalous.

```{r}

# Calculations for assault rates.
combined_assault <- combined_assault |>
  mutate(month_number = month(month)) |>
  group_by(month_number) |>
  mutate(
    mean_pg_assault = mean(pg_total),
    sd_pg_assault = sd(pg_total),
    z_pg_assault = (pg_total - mean_pg_assault) / sd_pg_assault, 
    mean_moco_assault = mean(moco_total),
    sd_moco_assault = sd(moco_total),
    z_moco_assault = (moco_total - mean_moco_assault) / sd_moco_assault
  ) |>
  ungroup()


# Calculations for automobile theft rates.
combined_autotheft <- combined_autotheft |>
  mutate(month_number = month(month)) |>
  group_by(month_number) |>
  mutate(
    mean_pg_autotheft = mean(pg_total),
    sd_pg_autotheft = sd(pg_total),
    z_pg_autotheft = (pg_total - mean_pg_autotheft) / sd_pg_autotheft, 
    mean_moco_autotheft = mean(moco_total),
    sd_moco_autotheft = sd(moco_total),
    z_moco_autotheft = (moco_total - mean_moco_autotheft) / sd_moco_autotheft
  ) |>
  ungroup()


# Calculations for burglary rates.
combined_burglary <- combined_burglary |>
  mutate(month_number = month(month)) |>
  group_by(month_number) |>
  mutate(
    mean_pg_burglary = mean(pg_total),
    sd_pg_burglary = sd(pg_total),
    z_pg_burglary = (pg_total - mean_pg_burglary) / sd_pg_burglary, 
    mean_moco_burglary = mean(moco_total),
    sd_moco_burglary = sd(moco_total),
    z_moco_burglary = (moco_total - mean_moco_burglary) / sd_moco_burglary
  ) |>
  ungroup()


# Calculations for homicide rates.
combined_homicide <- combined_homicide |>
  mutate(month_number = month(month)) |>
  group_by(month_number) |>
  mutate(
    mean_pg_homicide = mean(pg_total),
    sd_pg_homicide = sd(pg_total),
    z_pg_homicide = (pg_total - mean_pg_homicide) / sd_pg_homicide, 
    mean_moco_homicide = mean(moco_total),
    sd_moco_homicide = sd(moco_total),
    z_moco_homicide = (moco_total - mean_moco_homicide) / sd_moco_homicide
  ) |>
  ungroup()


# Calculations for robbery rates.
combined_robbery <- combined_robbery |>
  mutate(month_number = month(month)) |>
  group_by(month_number) |>
  mutate(
    mean_pg_robbery = mean(pg_total),
    sd_pg_robbery = sd(pg_total),
    z_pg_robbery = (pg_total - mean_pg_robbery) / sd_pg_robbery, 
    mean_moco_robbery = mean(moco_total),
    sd_moco_robbery = sd(moco_total),
    z_moco_robbery = (moco_total - mean_moco_robbery) / sd_moco_robbery
  ) |>
  ungroup()


# Calculations for sex offense rates.
combined_sexoffense <- combined_sexoffense |>
  mutate(month_number = month(month)) |>
  group_by(month_number) |>
  mutate(
    mean_pg_sexoffense = mean(pg_total),
    sd_pg_sexoffense = sd(pg_total),
    z_pg_sexoffense = (pg_total - mean_pg_sexoffense) / sd_pg_sexoffense, 
    mean_moco_sexoffense = mean(moco_total),
    sd_moco_sexoffense = sd(moco_total),
    z_moco_sexoffense = (moco_total - mean_moco_sexoffense) / sd_moco_sexoffense
  ) |>
  ungroup()


# Calculations for theft rates.
combined_theft <- combined_theft |>
  mutate(month_number = month(month)) |>
  group_by(month_number) |>
  mutate(
    mean_pg_theft = mean(pg_total),
    sd_pg_theft = sd(pg_total),
    z_pg_theft = (pg_total - mean_pg_theft) / sd_pg_theft, 
    mean_moco_theft = mean(moco_total),
    sd_moco_theft = sd(moco_total),
    z_moco_theft = (moco_total - mean_moco_theft) / sd_moco_theft
  ) |>
  ungroup()


# Calculations for theft from automobile rates.
combined_theftfromauto <- combined_theftfromauto |>
  mutate(month_number = month(month)) |>
  group_by(month_number) |>
  mutate(
    mean_pg_theftfromauto = mean(pg_total),
    sd_pg_theftfromauto = sd(pg_total),
    z_pg_theftfromauto = (pg_total - mean_pg_theftfromauto) / sd_pg_theftfromauto, 
    mean_moco_theftfromauto = mean(moco_total),
    sd_moco_theftfromauto = sd(moco_total),
    z_moco_theftfromauto = (moco_total - mean_moco_theftfromauto) / sd_moco_theftfromauto
  ) |>
  ungroup()
```



## Step 10: I filtered out unusual z-scores, meaning z-scores that had an absolute value of two or above. This would help me determine whether the apparent spikes and drops in crime rates, as displayed in the graphs, had actual meaning.

```{r}

# Filters for assault rates.
unusual_pg_assault <- combined_assault |>
  filter(abs(z_pg_assault) > 2) |>
  arrange(desc(abs(z_pg_assault)))

unusual_moco_assault <- combined_assault |>
  filter(abs(z_moco_assault) > 2) |>
  arrange(desc(abs(z_moco_assault)))



# Filters for automobile theft rates.
unusual_pg_autotheft <- combined_autotheft |>
  filter(abs(z_pg_autotheft) > 2) |>
  arrange(desc(abs(z_pg_autotheft)))

unusual_moco_autotheft <- combined_autotheft |>
  filter(abs(z_moco_autotheft) > 2) |>
  arrange(desc(abs(z_moco_autotheft)))



# Filters for burglary rates.
unusual_pg_burglary <- combined_burglary |>
  filter(abs(z_pg_burglary) > 2) |>
  arrange(desc(abs(z_pg_burglary)))

unusual_moco_burglary <- combined_burglary |>
  filter(abs(z_moco_burglary) > 2) |>
  arrange(desc(abs(z_moco_burglary)))



# Filters for homicide rates.
unusual_pg_homicide <- combined_homicide |>
  filter(abs(z_pg_homicide) > 2) |>
  arrange(desc(abs(z_pg_homicide)))

unusual_moco_homicide <- combined_homicide |>
  filter(abs(z_moco_homicide) > 2) |>
  arrange(desc(abs(z_moco_homicide)))



# Filters for robbery rates.
unusual_pg_robbery <- combined_robbery |>
  filter(abs(z_pg_robbery) > 2) |>
  arrange(desc(abs(z_pg_robbery)))

unusual_moco_robbery <- combined_robbery |>
  filter(abs(z_moco_robbery) > 2) |>
  arrange(desc(abs(z_moco_robbery)))



# Filters for sex offense rates.
unusual_pg_sexoffense <- combined_sexoffense |>
  filter(abs(z_pg_sexoffense) > 2) |>
  arrange(desc(abs(z_pg_sexoffense)))

unusual_moco_sexoffense <- combined_sexoffense |>
  filter(abs(z_moco_sexoffense) > 2) |>
  arrange(desc(abs(z_moco_sexoffense)))



# Filters for theft rates.
unusual_pg_theft <- combined_theft |>
  filter(abs(z_pg_theft) > 2) |>
  arrange(desc(abs(z_pg_theft)))

unusual_moco_theft <- combined_theft |>
  filter(abs(z_moco_theft) > 2) |>
  arrange(desc(abs(z_moco_theft)))



# Filters for theft from automobile rates.
unusual_pg_theftfromauto <- combined_theftfromauto |>
  filter(abs(z_pg_theftfromauto) > 2) |>
  arrange(desc(abs(z_pg_theftfromauto)))

unusual_moco_theftfromauto <- combined_theftfromauto |>
  filter(abs(z_moco_theftfromauto) > 2) |>
  arrange(desc(abs(z_moco_theftfromauto)))
```



## Step 11: I graphed the crimes again, highlighting the months with unusual z-scores in red. My graph titles summarized my findings.

```{r}

# Graph for Prince George's County assault rates.
ggplot(combined_assault, aes(x = month, y = pg_total, color = abs(z_pg_assault) > 2)) +
  geom_point(alpha = 0.5) +
  scale_color_manual(values = c("gray", "red"), 
                     name = "Unusual Value",
                     labels = c("Normal", "Unusual")) +
  theme_minimal() +
  labs(title = "Prince George's County, Maryland, had unusually low assault rate in October 2019",
       subtitle = "Unusual months have z-scores with absolute values > 2",
       x = "Month",
       y = "Assault rate per 10,000 residents",
       caption = "Source: Prince George's County crime archives")

# Graph for Montgomery County assault rates.
ggplot(combined_assault, aes(x = month, y = moco_total, color = abs(z_moco_assault) > 2)) +
  geom_point(alpha = 0.5) +
  scale_color_manual(values = c("gray", "red"), 
                     name = "Unusual Value",
                     labels = c("Normal", "Unusual")) +
  theme_minimal() +
  labs(title = "Montgomery County, Maryland, had unusually low assault rates in late 2020",
       subtitle = "Unusual months have z-scores with absolute values > 2",
       x = "Month",
       y = "Assault rate per 10,000 residents",
       caption = "Source: Montgomery County crime archives")


# Graph for Prince George's County automobile theft rates.
ggplot(combined_autotheft, aes(x = month, y = pg_total, color = abs(z_pg_autotheft) > 2)) +
  geom_point(alpha = 0.5) +
  scale_color_manual(values = c("gray", "red"), 
                     name = "Unusual Value",
                     labels = c("Normal", "Unusual")) +
  theme_minimal() +
  labs(title = "Prince George's County, Maryland, had unusually high auto theft rates for five months in 2023",
       subtitle = "Unusual months have z-scores with absolute values > 2",
       x = "Month",
       y = "Auto theft rate per 10,000 residents",
       caption = "Source: Prince George's County crime archives")


# Graph for Montgomery County automobile theft rates.
ggplot(combined_autotheft, aes(x = month, y = moco_total, color = abs(z_moco_autotheft) > 2)) +
  geom_point(alpha = 0.5) +
  scale_color_manual(values = c("gray", "red"), 
                     name = "Unusual Value",
                     labels = c("Normal", "Unusual")) +
  theme_minimal() +
  labs(title = "Montgomery County, Maryland, had unusually high auto theft rates in late 2023",
       subtitle = "Unusual months have z-scores with absolute values > 2",
       x = "Month",
       y = "Auto theft rate per 10,000 residents",
       caption = "Source: Montgomery County crime archives")


# Graph for Prince George's County burglary rates.
ggplot(combined_burglary, aes(x = month, y = pg_total, color = abs(z_pg_burglary) > 2)) +
  geom_point(alpha = 0.5) +
  scale_color_manual(values = c("gray", "red"), 
                     name = "Unusual Value",
                     labels = c("Normal", "Unusual")) +
  theme_minimal() +
  labs(title = "Prince George's County, Maryland, had unusually high burglary rates for six months in 2017 and early 2018",
       subtitle = "Unusual months have z-scores with absolute values > 2",
       x = "Month",
       y = "Burglary rate per 10,000 residents",
       caption = "Source: Prince George's County crime archives")


# Graph for Montgomery County burglary rates.
ggplot(combined_burglary, aes(x = month, y = moco_total, color = abs(z_moco_burglary) > 2)) +
  geom_point(alpha = 0.5) +
  scale_color_manual(values = c("gray", "red"), 
                     name = "Unusual Value",
                     labels = c("Normal", "Unusual")) +
  theme_minimal() +
  labs(title = "Montgomery County, Maryland, had four months with unusual burglary rates between 2017 and 2025",
       subtitle = "Unusual months have z-scores with absolute values > 2",
       x = "Month",
       y = "Burglary rate per 10,000 residents",
       caption = "Source: Montgomery County crime archives")


# Graph for Prince George's County homicide rates.
ggplot(combined_homicide, aes(x = month, y = pg_total, color = abs(z_pg_homicide) > 2)) +
  geom_point(alpha = 0.5) +
  scale_color_manual(values = c("gray", "red"), 
                     name = "Unusual Value",
                     labels = c("Normal", "Unusual")) +
  theme_minimal() +
  labs(title = "Prince George's County, Maryland, had unusually high homicide rates for three months between 2020 and 2023",
       subtitle = "Unusual months have z-scores with absolute values > 2",
       x = "Month",
       y = "Homicide rate per 10,000 residents",
       caption = "Source: Prince George's County crime archives")


# Graph for Montgomery County homicide rates.
ggplot(combined_homicide, aes(x = month, y = moco_total, color = abs(z_moco_homicide) > 2)) +
  geom_point(alpha = 0.5) +
  scale_color_manual(values = c("gray", "red"), 
                     name = "Unusual Value",
                     labels = c("Normal", "Unusual")) +
  theme_minimal() +
  labs(title = "Montgomery County, Maryland, had unusually high homicide rates for four months between 2017 and 2024",
       subtitle = "Unusual months have z-scores with absolute values > 2",
       x = "Month",
       y = "Homicide rate per 10,000 residents",
       caption = "Source: Montgomery County crime archives")


# Graph for Prince George's County robbery rates.
ggplot(combined_robbery, aes(x = month, y = pg_total, color = abs(z_pg_robbery) > 2)) +
  geom_point(alpha = 0.5) +
  scale_color_manual(values = c("gray", "red"), 
                     name = "Unusual Value",
                     labels = c("Normal", "Unusual")) +
  theme_minimal() +
  labs(title = "Prince George's County, Maryland, had unusually high robbery rates for three months between 2023 and 2024",
       subtitle = "Unusual months have z-scores with absolute values > 2",
       x = "Month",
       y = "Robbery rate per 10,000 residents",
       caption = "Source: Prince George's County crime archives")


# Graph for Montgomery County robbery rates.
ggplot(combined_robbery, aes(x = month, y = moco_total, color = abs(z_moco_robbery) > 2)) +
  geom_point(alpha = 0.5) +
  scale_color_manual(values = c("gray", "red"), 
                     name = "Unusual Value",
                     labels = c("Normal", "Unusual")) +
  theme_minimal() +
  labs(title = "Montgomery County, Maryland, had unusually high robbery rates for seven months between 2017 and 2024",
       subtitle = "Unusual months have z-scores with absolute values > 2",
       x = "Month",
       y = "Robbery rate per 10,000 residents",
       caption = "Source: Montgomery County crime archives")


# Graph for Prince George's County sex offense rates.
ggplot(combined_sexoffense, aes(x = month, y = pg_total, color = abs(z_pg_sexoffense) > 2)) +
  geom_point(alpha = 0.5) +
  scale_color_manual(values = c("gray", "red"), 
                     name = "Unusual Value",
                     labels = c("Normal", "Unusual")) +
  theme_minimal() +
  labs(title = "Prince George's County, Maryland, had unusually high sex offense rates for three months between 2018 and 2019",
       subtitle = "Unusual months have z-scores with absolute values > 2",
       x = "Month",
       y = "Sex offense rate per 10,000 residents",
       caption = "Source: Prince George's County crime archives")


# Graph for Montgomery County sex offense rates.
ggplot(combined_sexoffense, aes(x = month, y = moco_total, color = abs(z_moco_sexoffense) > 2)) +
  geom_point(alpha = 0.5) +
  scale_color_manual(values = c("gray", "red"), 
                     name = "Unusual Value",
                     labels = c("Normal", "Unusual")) +
  theme_minimal() +
  labs(title = "Montgomery County, Maryland, had no months with unusual sex offense rates between April 2017 and March 2026",
       subtitle = "Unusual months have z-scores with absolute values > 2",
       x = "Month",
       y = "Sex offense rate per 10,000 residents",
       caption = "Source: Montgomery County crime archives")


# Graph for Prince George's County theft rates.
ggplot(combined_theft, aes(x = month, y = pg_total, color = abs(z_pg_theft) > 2)) +
  geom_point(alpha = 0.5) +
  scale_color_manual(values = c("gray", "red"), 
                     name = "Unusual Value",
                     labels = c("Normal", "Unusual")) +
  theme_minimal() +
  labs(title = "Prince George's County, Maryland, had unusually low theft rates for two months in 2021",
       subtitle = "Unusual months have z-scores with absolute values > 2",
       x = "Month",
       y = "Theft rate per 10,000 residents",
       caption = "Source: Prince George's County crime archives")


# Graph for Montgomery County theft rates.
ggplot(combined_theft, aes(x = month, y = moco_total, color = abs(z_moco_theft) > 2)) +
  geom_point(alpha = 0.5) +
  scale_color_manual(values = c("gray", "red"), 
                     name = "Unusual Value",
                     labels = c("Normal", "Unusual")) +
  theme_minimal() +
  labs(title = "Montgomery County, Maryland, had unusually low theft rates for two months between 2023 and 2024",
       subtitle = "Unusual months have z-scores with absolute values > 2",
       x = "Month",
       y = "Theft rate per 10,000 residents",
       caption = "Source: Montgomery County crime archives")


# Graph for Prince George's County theft from automobile rates.
ggplot(combined_theftfromauto, aes(x = month, y = pg_total, color = abs(z_pg_theftfromauto) > 2)) +
  geom_point(alpha = 0.5) +
  scale_color_manual(values = c("gray", "red"), 
                     name = "Unusual Value",
                     labels = c("Normal", "Unusual")) +
  theme_minimal() +
  labs(title = "Prince George's County, Maryland, had unusual theft from automobile rates for two months between 2021 and 2023",
       subtitle = "Unusual months have z-scores with absolute values > 2",
       x = "Month",
       y = "Theft from automobile rate per 10,000 residents",
       caption = "Source: Prince George's County crime archives")


# Graph for Montgomery County theft from automobile rates.
ggplot(combined_theftfromauto, aes(x = month, y = moco_total, color = abs(z_moco_theftfromauto) > 2)) +
  geom_point(alpha = 0.5) +
  scale_color_manual(values = c("gray", "red"), 
                     name = "Unusual Value",
                     labels = c("Normal", "Unusual")) +
  theme_minimal() +
  labs(title = "Montgomery County, Maryland, had unusual theft from automobile rates for two months between 2018 and 2020",
       subtitle = "Unusual months have z-scores with absolute values > 2",
       x = "Month",
       y = "Theft from automobile rate per 10,000 residents",
       caption = "Source: Montgomery County crime archives")
```

## Step 11 results:

## The spike in automobile thefts in both counties that I noticed in the graph was corroborated by my z-score analysis, further proving that said spike was unusual and underlining the need for additional research. 

## The anomalies in burglary rates in Prince George's County caught my eye because the graphs appeared to show an overall decrease. I will need to research this further to detect the causes of these anomalies.

## The spike in homicide rates in Prince George's County that I noticed in the graph is corroborated by my z-score analysis, indicating a need for further research. 

## The spike in robbery rates in Prince George's County that I noticed in the graph was corroborated by my z-score analysis, indicating a need for further research. 

## The spike in sex offense rates in Prince George's County that I noticed in the graph was corroborated by my z-score analysis, indicating a need for further research. 

## The spike in theft rates in both counties that I noticed in the graph was not corroborated by my z-score analysis. However, those spikes may still be significant, and I need to research them.




## Step 12: I renamed the columns in my eight large dataframes so I could combine them for further analysis. 

```{r}
rename_rates <- function(combined_assault, crime_name) {
  combined_assault |>
    rename_with(~ "pg_rate", starts_with("pg_") & ends_with("_rate")) |>
    rename_with(~ "moco_rate", starts_with("moco_") & ends_with("_rate"))}

rename_rates <- function(combined_autotheft, crime_name) {
  combined_autotheft |>
    rename_with(~ "pg_rate", starts_with("pg_") & ends_with("_rate")) |>
    rename_with(~ "moco_rate", starts_with("moco_") & ends_with("_rate"))}

rename_rates <- function(combined_burglary, crime_name) {
  combined_burglary |>
    rename_with(~ "pg_rate", starts_with("pg_") & ends_with("_rate")) |>
    rename_with(~ "moco_rate", starts_with("moco_") & ends_with("_rate"))}

rename_rates <- function(combined_homicide, crime_name) {
  combined_homicide |>
    rename_with(~ "pg_rate", starts_with("pg_") & ends_with("_rate")) |>
    rename_with(~ "moco_rate", starts_with("moco_") & ends_with("_rate"))}

rename_rates <- function(combined_robbery, crime_name) {
  combined_robbery |>
    rename_with(~ "pg_rate", starts_with("pg_") & ends_with("_rate")) |>
    rename_with(~ "moco_rate", starts_with("moco_") & ends_with("_rate"))}

rename_rates <- function(combined_sexoffense, crime_name) {
  combined_sexoffense |>
    rename_with(~ "pg_rate", starts_with("pg_") & ends_with("_rate")) |>
    rename_with(~ "moco_rate", starts_with("moco_") & ends_with("_rate"))}

rename_rates <- function(combined_theft, crime_name) {
  combined_theft |>
    rename_with(~ "pg_rate", starts_with("pg_") & ends_with("_rate")) |>
    rename_with(~ "moco_rate", starts_with("moco_") & ends_with("_rate"))}

rename_rates <- function(combined_theftfromauto, crime_name) {
  combined_theftfromauto |>
    rename_with(~ "pg_rate", starts_with("pg_") & ends_with("_rate")) |>
    rename_with(~ "moco_rate", starts_with("moco_") & ends_with("_rate"))}


combined_all <- bind_rows(
  rename_rates(combined_assault),
  rename_rates(combined_autotheft),
  rename_rates(combined_burglary),
  rename_rates(combined_homicide),
  rename_rates(combined_robbery),
  rename_rates(combined_sexoffense),
  rename_rates(combined_theft),
  rename_rates(combined_theftfromauto)
)
```



## Step 13: Then, I ran a correlation test to see if crime rates were correlated between both counties.

```{r}
combined_all |>
  group_by(category) |>
  summarise(correlation = cor(pg_rate, moco_rate))
```

## Step 13 results: The result that fascinated me the most was the high correlation coefficient between automobile theft rates in both counties. Though I noticed that the rates between the counties almost mirrored each other in the graph, I forgot to mention it. However, when thought about in the context of the real world, it makes sense. To an extent, neighboring counties should experience similar increases and decreases in crime rates. Based on this viewpoint, the lack of correlation between any other crime category is more surprising.




## Step 14: I rearranged Prince George's County's crime data, then calculated the correlation coefficients between individual crimes. 

```{r}
# Rearranged data for Prince George's County.
pg_wide <- combined_all |>
  select(month, category, pg_rate) |>
  pivot_wider(names_from = category, values_from = pg_rate)


cor(pg_wide$`assault`, pg_wide$`auto theft`)

cor(pg_wide$`assault`, pg_wide$`burglary`)

cor(pg_wide$`assault`, pg_wide$`robbery`)

cor(pg_wide$`assault`, pg_wide$`sex offense`)

cor(pg_wide$`assault`, pg_wide$`theft`)

cor(pg_wide$`assault`, pg_wide$`theft from auto`)


cor(pg_wide$`auto theft`, pg_wide$`burglary`)

cor(pg_wide$`auto theft`, pg_wide$`robbery`)

cor(pg_wide$`auto theft`, pg_wide$`sex offense`)

cor(pg_wide$`auto theft`, pg_wide$`theft`)

cor(pg_wide$`auto theft`, pg_wide$`theft from auto`)


cor(pg_wide$`burglary`, pg_wide$`robbery`)

cor(pg_wide$`burglary`, pg_wide$`sex offense`)

cor(pg_wide$`burglary`, pg_wide$`theft`)

cor(pg_wide$`burglary`, pg_wide$`theft from auto`)


cor(pg_wide$`robbery`, pg_wide$`sex offense`)

cor(pg_wide$`robbery`, pg_wide$`theft`)

cor(pg_wide$`robbery`, pg_wide$`theft from auto`)


cor(pg_wide$`sex offense`, pg_wide$`theft`)

cor(pg_wide$`sex offense`, pg_wide$`theft from auto`)


cor(pg_wide$`theft`, pg_wide$`theft from auto`)
```
## Step 14 results: I found that none of the crimes were strongly correlated with each other on their own, which did not surprise me, as the nature and motivations behind different crimes are different. Unfortunately, when the data was rearranged, the homicide data was messed up. Consequently, it was excluded from my analysis.




## Step 15: I then created multiple regression models for each crime category. Below are the models I deemed most successful (they had relatively high R-squared values). 

```{r}

# A model attempting to predict theft based on automobile theft, theft from automobiles, burglary and robbery.

pg_theft_model <- lm(theft ~ `auto theft` + `theft from auto` + burglary + robbery, data = pg_wide)
summary(pg_theft_model)


# A model attempting to predict assault based on sex offenses, theft, burglary and robbery.

pg_assault_model <- lm(assault ~ `sex offense` + theft + burglary + robbery, data = pg_wide)
summary(pg_assault_model)


# A model attempting to predict automobile theft based on theft, theft from automobiles, burglary and robbery.

pg_autotheft_model <- lm(`auto theft` ~ theft + `theft from auto` + burglary + robbery, data = pg_wide)
summary(pg_autotheft_model)


# A model attempting to predict burglary based on robbery, sex offenses, theft and theft from automobiles.

pg_burglary_model <- lm(burglary ~ robbery + `sex offense` + theft + `theft from auto`, data = pg_wide)
summary(pg_burglary_model)


# A model attempting to predict robbery based on assault, sex offenses, theft and theft from automobiles.

pg_robbery_model <- lm(robbery ~ assault + `sex offense` + theft + `theft from auto`, data = pg_wide)
summary(pg_robbery_model)


# A model attempting to predict sex offenses based on robbery, burglary and assault.

pg_sexoffense_model <- lm(`sex offense` ~ robbery + burglary + assault, data = pg_wide)
summary(pg_sexoffense_model)


# A model attempting to predict theft from automobiles based on theft, automobile theft, assault and burglary.

pg_theftfromauto_model <- lm(`theft from auto` ~ theft + `auto theft` + assault + burglary + robbery, data = pg_wide)
summary(pg_theftfromauto_model)
```
## Step 15 results: I found that, despite the multiple variations I created, none of my models were very successful. However, the low correlation coefficients between individual crime categories made this result unsurprising. Unfortunately, the homicide data was messed up again, and consequently excluded from my analysis.




## Step 16: I rearranged Montgomery County's crime data, then calculated the correlation coefficients between individual crimes. 

```{r}

# Rearranged data for Montgomery County.
moco_wide <- combined_all |>
  select(month, category, moco_rate) |>
  pivot_wider(names_from = category, values_from = moco_rate)


cor(moco_wide$`assault`, moco_wide$`auto theft`)

cor(moco_wide$`assault`, moco_wide$`burglary`)

cor(moco_wide$`assault`, moco_wide$`robbery`)

cor(moco_wide$`assault`, moco_wide$`sex offense`)

cor(moco_wide$`assault`, moco_wide$`theft`)

cor(moco_wide$`assault`, moco_wide$`theft from auto`)


cor(moco_wide$`auto theft`, moco_wide$`burglary`)

cor(moco_wide$`auto theft`, moco_wide$`robbery`)

cor(moco_wide$`auto theft`, moco_wide$`sex offense`)

cor(moco_wide$`auto theft`, moco_wide$`theft`)

cor(moco_wide$`auto theft`, moco_wide$`theft from auto`)


cor(moco_wide$`burglary`, moco_wide$`robbery`)

cor(moco_wide$`burglary`, moco_wide$`sex offense`)

cor(moco_wide$`burglary`, moco_wide$`theft`)

cor(moco_wide$`burglary`, moco_wide$`theft from auto`)


cor(moco_wide$`robbery`, moco_wide$`sex offense`)

cor(moco_wide$`robbery`, moco_wide$`theft`)

cor(moco_wide$`robbery`, moco_wide$`theft from auto`)


cor(moco_wide$`sex offense`, moco_wide$`theft`)

cor(moco_wide$`sex offense`, moco_wide$`theft from auto`)


cor(moco_wide$`theft`, moco_wide$`theft from auto`)
```
## Step 16 results: I again found that none of the crimes were strongly correlated with each other on their own. Once again, the homicide data had errors and was excluded from my analysis.

## However, for both Prince George's and Montgomery counties, automobile theft and robbery had the highest correlation coefficients (approximately 0.58 and 0.67, respectively), which is interesting and may warrant further investigation.




## Step 17: I created multiple regression models for each crime category. Below are my most successful models. Once again, my models were relatively unsuccessful.

```{r}
# A model attempting to predict theft based on automobile theft, theft from automobiles, assault and burglary.

moco_theft_model <- lm(theft ~ `auto theft` + `theft from auto` + assault + burglary + robbery, data = moco_wide)
summary(moco_theft_model)


# A model attempting to predict assault based on theft, sex offenses, burglary and robbery.

moco_assault_model <- lm(assault ~ theft + `sex offense` + burglary + robbery, data = moco_wide)
summary(moco_assault_model)


# A model attempting to predict automobile theft based on theft, theft from automobiles, burglary and robbery.

moco_autotheft_model <- lm(`auto theft` ~ theft + `theft from auto` + burglary + robbery, data = moco_wide)
summary(moco_autotheft_model)


# A model attempting to predict burglary based on robbery, assault and theft.

moco_burglary_model <- lm(burglary ~ robbery + assault + theft, data = moco_wide)
summary(moco_burglary_model)


# A model attempting to predict robbery based on assault, theft and burglary.

moco_robbery_model <- lm(robbery ~ assault + theft + burglary, data = moco_wide)
summary(moco_robbery_model)


# A model attempting to predict sex offenses based on robbery, assault and theft.

moco_sexoffense_model <- lm(`sex offense` ~ robbery + assault + theft, data = moco_wide)
summary(moco_sexoffense_model)


# A model attempting to predict theft from automobiles based on theft and automobile theft.

moco_theftfromauto_model <- lm(`theft from auto` ~ theft + `auto theft`, data = moco_wide)
summary(moco_theftfromauto_model)
```
Hypothesis reevaluation :

The burglary rate for both counties fluctuated from April 2017 to 2025, but since then, Montgomery County's automobile theft has been higher, proving that part of my hypothesis. Prince George's County had higher rates of automobile theft than Montgomery County from April 2017 to March 2026, disproving that part of my hypothesis. The theft from automobile rate for both counties fluctuated during that time period, meaning that part of my hypothesis cannot be proven.

Montgomery County had higher assault rates than Prince George's County from April 2017 to March 2026, disproving that part of my hypothesis. Montgomery County generally had higher rates of sex offenses than Prince George's County during that same time period, disproving another part of my hypothesis. Montgomery County had higher theft rates than Prince George's County during that same time period, disproving another part of my hypothesis.

Additionally, Prince George's County generally had higher homicide and robbery rates than Montgomery County. 

I did not calculate overall crime rates based on the eight categories I selected because Montgomery County's high assault rates likely would have skewed those results. Additionally, both counties include and exclude different crime categories. For example, Prince George's County does not report financial crimes or crimes related to unlawful substances or alcohol to the public, while Montgomery County does. On the other hand, it reports car accidents, while Montgomery County does not. Consequently, I thought it best to only analyze the eight crime categories they had in common. This means that I decided to ignore my third hypothesis.


Story pitch:

The main finding from my analysis is that Montgomery County had an assault rate approximately three times higher than Prince George's County from April 2017 to March 2026. If I were to pitch a story based on my analysis, I would base it on that finding, with a focus on investigating the cause of that discrepancy. 

One problem with this finding is that it may have been skewed by my categorization of assault. I combined three of Prince George's County's crime categories into an "assault" category, while I combined 15 for Montgomery County. I may have accidentally miscategorized one of these specific categories, or Montgomery County may have counted crimes for multiple categories, such as counting sexual assault as assault. To rectify this issue, I would need to have my categorizations reevaluated by someone else. I would also need to contact the Montgomery and Prince George's counties’ police departments to ask them about their crime categorizations.

Another problem with this finding is that the per capita rates for both counties were based on 2022 population data. This may have slightly skewed the assault rates. To rectify this issue, I would need to edit my dataframes to add population estimates from 2017 to 2026.

To report on this finding, I would need to take many steps. I would need to have my categorizations reevaluated to ensure their accuracy, and add incremental population data for both Montgomery and Prince George's counties.

I would need to contact the Montgomery County Police Department to discuss its categorizations of assault and the increase in assaults in 2023 and 2024. I would also need to evaluate if this increase was statistically significant, and if so, evaluate its effect size.

I would need to contact the Prince George's County Police Department to discuss its categorization of assault and ask about the unusual dip in assaults in October 2019. I would also send a public records request for its financial, unlawful substance-related and alcohol-related crime records from April 2017 to March 2026. With that information, I could calculate the overall crime rate for both counties and evaluate if their assault rates can predict their overall crime rates.

I would need to look at local news archives and other sources of information to investigate any documented spikes or dips in assault rates, then see if they are reflected in the data. I would want to analyze national assault rates to see if both counties' rates were unusually high or low.

I would need to incorporate demographic information, such as median income, race and education level, into my dataframes and create regression models to evaluate if these factors could predict assault rates in either Montgomery or Prince George's County. I would also want to map out the locations of each crime and look for patterns in crime locations. Additionally, I would want to interview residents of areas with high assault rates and ask them about their perception of assault in their communities.
