---
title: "final_project"
author: Josephine Andoh
output: html_document
---


## I loaded the tidyverse, knitr, lubridate, dplyr and scales so I could analyze and graph my datasets.

```{r}
library(tidyverse)

knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(knitr) 

library(lubridate)

library(dplyr)

library(scales)
```



## I loaded crime data from Prince George's County from 2017 to 2023 and 2023 to 2026, then combined those datasets to facilitate my analysis. I then loaded crime data from Montgomery County from 2017 to 2026.

```{r}
pgcrime17_23 <- read_csv("PGCRIME17-23.csv")

pgcrime23_26 <- read_csv("PGCRIME23-26.csv")

mococrime <- read_csv("MOCOCRIME.zip")

pgcrime <- bind_rows(pgcrime17_23, pgcrime23_26)
```



## I sorted all the crimes into seven categories (assault, burglary, auto theft, theft from auto, robbery, sex offenses and homicide) to facilitate my analysis.

```{r}
pgcrime <- pgcrime |> 
  mutate(category = case_when(
    str_detect(`Clearance Code Inc Type`, "ASSAULT") ~ "assault",
    str_detect(`Clearance Code Inc Type`, "ASSAULT, WEAPON") ~ "assault",
    str_detect(`Clearance Code Inc Type`, "ASSAULT, SHOOTING") ~ "assault",
    str_detect(`Clearance Code Inc Type`, "B & E, RESIDENTIAL") ~ "burglary",
    str_detect(`Clearance Code Inc Type`, "B & E, RESIDENTIAL (VACANT)") ~ "burglary",
    str_detect(`Clearance Code Inc Type`, "B & E, COMMERCIAL") ~ "burglary",
    str_detect(`Clearance Code Inc Type`, "B & E, OTHER") ~ "burglary",
    str_detect(`Clearance Code Inc Type`, "AUTO, STOLEN") ~ "auto theft",
    str_detect(`Clearance Code Inc Type`, "AUTO, STOLEN & RECOVERED") ~ "auto theft",
    str_detect(`Clearance Code Inc Type`, "ROBBERY, OTHER") ~ "robbery",
    str_detect(`Clearance Code Inc Type`, "ROBBERY, COMMERCIAL") ~ "robbery",
    str_detect(`Clearance Code Inc Type`, "ROBBERY, RESIDENTIAL") ~ "robbery",
    str_detect(`Clearance Code Inc Type`, "ROBBERY, VEHICLE") ~ "robbery",
    str_detect(`Clearance Code Inc Type`, "ROBBERY, VEHICLE") ~ "robbery",
    str_detect(`Clearance Code Inc Type`, "THEFT FROM AUTO") ~ "theft from auto",
    str_detect(`Clearance Code Inc Type`, "THEFT") ~ "theft",
    str_detect(`Clearance Code Inc Type`, "SEX OFFENSE") ~ "sex offense",
    str_detect(`Clearance Code Inc Type`, "HOMICIDE") ~ "homicide",
    .default = NA
  )
)


mococrime <- mococrime |>
  mutate(category = case_when(
    str_detect(`Crime Name3`, "ASSAULT - 2ND DEGREE") ~ "assault",
    str_detect(`Crime Name3`, "ASSAULT - INTIMIDATION (INCLUDES STALKING)") ~ "assault",
    str_detect(`Crime Name3`, "ASSAULT - AGGRAVATED - PUB OFF-OTHER WEAPON") ~ "assault",
    str_detect(`Crime Name3`, "ASSAULT - AGGRAVATED - POL OFF-OTHER WEAPON") ~ "assault",
    str_detect(`Crime Name3`, "ASSAULT - AGGRAVATED - FAMILY-GUN") ~ "assault",
    str_detect(`Crime Name3`, "ASSAULT - SIMPLE") ~ "assault",
    str_detect(`Crime Name3`, "ASSAULT - AGGRAVATED - GUN") ~ "assault",
    str_detect(`Crime Name3`, "ASSAULT - AGGRAVATED - NON-FAMILY-OTHER WEAPON") ~ "assault",
    str_detect(`Crime Name3`, "ASSAULT - AGGRAVATED - FAMILY-STRONG-ARM") ~ "assault",
    str_detect(`Crime Name3`, "ASSAULT - AGGRAVATED - NON-FAMILY-GUN") ~ "assault",
    str_detect(`Crime Name3`, "ASSAULT - AGGRAVATED - NON-FAMILY-STRONG-ARM") ~ "assault",
    str_detect(`Crime Name3`, "ATTEMPTED ASSAULT AGGRAVATED POL OFF KNIFE") ~ "assault",
    str_detect(`Crime Name3`, "ASSAULT - AGGRAVATED - POL OFF-STRONG-ARM") ~ "assault",
    str_detect(`Crime Name3`, "ASSAULT - AGGRAVATED - POL OFF-GUN") ~ "assault",
    str_detect(`Crime Name3`, "ASSAULT - AGGRAVATED - PUB OFF-STRONG-ARM") ~ "assault",
    str_detect(`Crime Name3`, "ASSAULT - AGGRAVATED - PUB OFF-GUN") ~ "assault",
    
    .default = NA
  )
)

mococrime <- mococrime |>
  mutate(category = case_when(
    str_detect(`Crime Name3`, "SEX OFFENSE - AGAINST CHILD-FONDLING") ~ "sex offense",
    str_detect(`Crime Name3`, "SEX ASSAULT - FONDLING") ~ "sex offense",
    str_detect(`Crime Name3`, "SEX OFFENSE - PEEPING TOM") ~ "sex offense",
    str_detect(`Crime Name3`, "SEX OFFENSE - INDECENT EXPOSURE") ~ "sex offense",
    str_detect(`Crime Name3`, "COMM SEX OFF - KEEPING HOUSE ILL FAME") ~ "sex offense",
    str_detect(`Crime Name3`, "SEX OFFENSE - INCEST WITH ADULT") ~ "sex offense",
    str_detect(`Crime Name3`, "SEX ASSAULT - SODOMY-GIRL-GUN") ~ "sex offense",
    str_detect(`Crime Name3`, "SEX ASSAULT - SODOMY-GIRL-WEAPON") ~ "sex offense",
    str_detect(`Crime Name3`, "COMM SEX OFF - PROCURE FOR PROSTITUTE (PIMPING)") ~ "sex offense",
    str_detect(`Crime Name3`, "COMM SEX OFF - PROCURE PROSTITUTE - MINOR") ~ "sex offense",
    str_detect(`Crime Name3`, "SEX ASSAULT - SODOMY-WOMAN-GUN") ~ "sex offense",
    str_detect(`Crime Name3`, "COMM SEX OFF - PROSTITUTION") ~ "sex offense",
    str_detect(`Crime Name3`, "HOMOSEXUAL ACT WITH BOY") ~ "sex offense",
    str_detect(`Crime Name3`, "SEX OFFENSE - INCEST WITH MINOR") ~ "sex offense",
    str_detect(`Crime Name3`, "SEX ASSAULT - SODOMY-MAN-WEAPON") ~ "sex offense",
    str_detect(`Crime Name3`, "COMM SEX OFF - HOMOSEXUAL PROSTITUTION") ~ "sex offense",
    str_detect(`Crime Name3`, "SEX ASSAULT - SODOMY-WOMAN-WEAPON") ~ "sex offense",
    str_detect(`Crime Name3`, "SEX ASSAULT - SODOMY-MAN-GUN") ~ "sex offense",
    str_detect(`Crime Name3`, "COMM SEX OFF - TRANSPORT FEMALE INTERSTATE FOR IMM") ~ "sex offense",
    .default = category
  )
)

mococrime <- mococrime |>
  mutate(category = case_when(
    str_detect(`Crime Name3`, "ROBBERY - STREET-GUN") ~ "robbery",
    str_detect(`Crime Name3`, "ROBBERY - RESIDENTIAL-STRONG-ARM") ~ "robbery",
    str_detect(`Crime Name3`, "ROBBERY - STREET-STRONG-ARM") ~ "robbery",
    str_detect(`Crime Name3`, "ROBBERY - STREET-OTHER WEAPON") ~ "robbery",
    str_detect(`Crime Name3`, "ROBBERY - BUSINESS-GUN") ~ "robbery",
    str_detect(`Crime Name3`, "ROBBERY - FORCIBLE PURSE SNATCHING") ~ "robbery",
    str_detect(`Crime Name3`, "ROBBERY (DESCRIBE OFFENSE)") ~ "robbery",
    str_detect(`Crime Name3`, "ROBBERY - RESIDENTIAL-GUN") ~ "robbery",
    str_detect(`Crime Name3`, "ROBBERY - BUSINESS-OTHER WEAPON") ~ "robbery",
    str_detect(`Crime Name3`, "ROBBERY - CARJACKING - ARMED") ~ "robbery",
    str_detect(`Crime Name3`, "ROBBERY - RESIDENTIAL-OTHER WEAPON") ~ "robbery",
    str_detect(`Crime Name3`, "ROBBERY - BANKING-TYPE INST") ~ "robbery",
    .default = category
  )
)

mococrime <- mococrime |>
  mutate(category = case_when(
    str_detect(`Crime Name3`, "BURGLARY - SAFE-VAULT") ~ "burglary",
    str_detect(`Crime Name3`, "BURGLARY - BANKING-TYPE INST") ~ "burglary",
    str_detect(`Crime Name3`, "BURGLARY (DESCRIBE OFFENSE)") ~ "burglary",
    str_detect(`Crime Name3`, "BURGLARY - NO FORCED ENTRY-NONRESIDENTIAL") ~ "burglary",
    str_detect(`Crime Name3`, "BURGLARY - FORCED ENTRY-RESIDENTIAL") ~ "burglary",
    str_detect(`Crime Name3`, "BURGLARY - FORCED ENTRY-NONRESIDENTIAL") ~ "burglary",
    str_detect(`Crime Name3`, "BURGLARY - NO FORCED ENTRY-RESIDENTIAL") ~ "burglary",
    .default = category
  )
)

mococrime <- mococrime |>
  mutate(category = case_when(
    str_detect(`Crime Name3`, "STOLEN VEHICLE (DESCRIBE OFFENSE)") ~ "auto theft",
    str_detect(`Crime Name3`, "AUTO THEFT - VEHICLE THEFT") ~ "auto theft",
    str_detect(`Crime Name3`, "AUTO THEFT - THEFT VEHICLE BY BAILEE") ~ "auto theft",
    str_detect(`Crime Name3`, "AUTO THEFT - STRIP STOLEN VEHICLE") ~ "auto theft",
    str_detect(`Crime Name3`, "AUTO THEFT - THEFT AND STRIP VEHICLE") ~ "auto theft",
    str_detect(`Crime Name3`, "AUTO THEFT - THEFT AND SALE VEHICLE") ~ "auto theft",
    str_detect(`Crime Name3`, "AUTO THEFT - THEFT AND USE VEHICLE OTHER CRIME") ~ "auto theft",
    .default = category
  )
)

mococrime <- mococrime |>
  mutate(category = case_when(
    str_detect(`Crime Name3`, "LARCENY - FROM AUTO") ~ "theft from auto",
    str_detect(`Crime Name3`, "LARCENY - AUTO PARTS") ~ "theft from auto",
    .default = category
  )
)

mococrime <- mococrime |>
  mutate(category = case_when(
    str_detect(`Crime Name3`, "HOMICIDE - WILLFUL KILL-GUN") ~ "homicide",
    str_detect(`Crime Name3`, "HOMICIDE (DESCRIBE OFFENSE)") ~ "homicide",
    str_detect(`Crime Name3`, "HOMICIDE - WILLFUL KILL") ~ "homicide",
    str_detect(`Crime Name3`, "HOMICIDE - WILLFUL KILL-FAMILY") ~ "homicide",
    str_detect(`Crime Name3`, "HOMICIDE - JUSTIFIABLE") ~ "homicide",
    str_detect(`Crime Name3`, "HOMICIDE - WILLFUL KILL-FAMILY-GUN") ~ "homicide",
    str_detect(`Crime Name3`, "HOMICIDE - WILLFUL KILL-NON-FAMILY-GUN") ~ "homicide",
    str_detect(`Crime Name3`, "HOMICIDE - WILLFUL KILL-NON-FAMILY") ~ "homicide",
    str_detect(`Crime Name3`, "HOMICIDE - NEGLIGENT MANSLAUGHTER") ~ "homicide",
    str_detect(`Crime Name3`, "HOMICIDE - BY NON-POLICE - JUSTIFIABLE") ~ "homicide",
    str_detect(`Crime Name3`, "HOMICIDE - JOHN OR JANE DOE - NO WARRANT") ~ "homicide",
    .default = category
  )
)

mococrime <- mococrime |>
  mutate(category = case_when(
    str_detect(`Crime Name3`, "LARCENY (DESCRIBE OFFENSE)") ~ "theft",
    str_detect(`Crime Name3`, "LARCENY - POCKET PICKING") ~ "theft",
    str_detect(`Crime Name3`, "LARCENY - FROM INTERSTATE SHIPMENT") ~ "theft",
    str_detect(`Crime Name3`, "LARCENY - THEFT OF US GOVERNMENT PROPERTY") ~ "theft",
    str_detect(`Crime Name3`, "LARCENY - FROM COIN MACHINE") ~ "theft",
    str_detect(`Crime Name3`, "LARCENY - FROM SHIPMENT") ~ "theft",
    str_detect(`Crime Name3`, "LARCENY - FROM BANKING-TYPE INST") ~ "theft",
    str_detect(`Crime Name3`, "LARCENY - FROM MALLS") ~ "theft",
    str_detect(`Crime Name3`, "LARCENY - FROM YARDS") ~ "theft",
    str_detect(`Crime Name3`, "LARCENY - PURSE SNATCHING - NO FORCE") ~ "theft",
    str_detect(`Crime Name3`, "LARCENY - POCKET PICKING") ~ "theft",
    str_detect(`Crime Name3`, "LARCENY - POSTAL") ~ "theft",
    str_detect(`Crime Name3`, "LARCENY - FROM BLDG") ~ "theft",
    str_detect(`Crime Name3`, "LARCENY - SHOPLIFTING") ~ "theft",
    .default = category
  )
)
```



## I filtered my data on a monthly basis so that I was not comparing too many individual data points.

```{r}
pgcrime <- pgcrime |>
  mutate(Date = mdy(Date)) |> 
  mutate(
    month = floor_date(Date, "month")
  )
```

```{r}
mococrime <- mococrime |> 
  mutate(Date = date(mdy_hm(Start_Date_Time)))|>
      mutate(month = floor_date(Date, "month"))
```



## I filtered my data from April 1, 2017, to April 1, 2026, so that I could compare crime rates over the same time period.

```{r}
new_pgcrime <- pgcrime |> filter(Date < '2026-04-01', Date > '2017-03-31')
```

```{r}
new_mococrime <- mococrime |> filter(Date < '2026-04-01', Date > '2017-03-31')
```



## I incorporated 2022 population data from both counties so that I could calculate per capita rates. I chose 2022 because it was close to the center of my range (2017 to 2026) and would most accurately reflect the population despite the COVID-19 pandemic. 

## I retrieved the Montgomery County data from this link (https://montgomeryplanning.org/blog-design/2024/06/montgomery-countys-population-rebound-part-1-recovering-from-the-covid-19-years/) and the Prince George's County data from this link (https://www.princegeorgescountymd.gov/sites/default/files/media-document/County%20Profile_0.pdf).

```{r}
new_mococrime <- new_mococrime %>%
  mutate(population = 1053067)

new_pgcrime <- new_pgcrime %>%
  mutate(population = 946971)
```


mean_assault = mean(total),
    sd_accidents = sd(total),
    z_score = (total - mean_accidents) / sd_accidents

unusual_days <- accidents |>
  filter(abs(z_score) > 2) |>
  arrange(desc(abs(z_score)))
  
cat("Number of days with unusual accident counts:", nrow(unusual_days), 
    "out of", nrow(accidents), "total days (", 
    round(nrow(unusual_days)/nrow(accidents)*100, 2), "%)")
    
unusual_days |>
  head(10) |>
  select(date, day_of_week, total, z_score) |>
  kable(digits = 2)
  
  
ggplot(accidents, aes(x = date, y = total, color = abs(z_score) > 2)) +
  geom_point(alpha = 0.5) +
  scale_color_manual(values = c("gray", "red"), 
                     name = "Unusual Value",
                     labels = c("Normal", "Unusual")) +
  theme_minimal() +
  labs(title = "Daily Accidents with Unusual Days Highlighted",
       subtitle = "Unusual days have z-scores with absolute values > 2",
       x = "Date",
       y = "Number of Accidents",
       caption = "Source: Prince George's County Accident Data")
       
       
extreme_outliers <- accidents |>
  filter(abs(z_score) > 3) |>
  arrange(desc(abs(z_score)))
  
  
  
if(nrow(extreme_outliers) > 0) {
  extreme_outliers |>
    select(date, day_of_week, total, z_score) |>
    kable(digits = 2)
} else {
  cat("No days with z-scores beyond ±3 were found in the dataset.")


## I calculated the number of specific crimes that occured in both counties per month, then calcuated the rate of said crimes per month per capita. I then combined the Prince George's and Montgomery counties' crime data to facilitate my comparison. 

```{r}
pg_assault <- new_pgcrime |> 
  filter(category == "assault") |> 
  group_by(category, month, population) |> 
  summarise(pg_total = n()) |> 
  mutate(pg_assault_rate = pg_total/population * 100000)

moco_assault <- new_mococrime |> 
  filter(category == "assault") |> 
  group_by(category, month, population) |> 
  summarise(moco_total = n()) |> 
  mutate(moco_assault_rate = moco_total/population * 100000)

combined_assault <- pg_assault |> 
  inner_join(moco_assault, join_by(category, month))



pg_sexoffense <- new_pgcrime |> 
  filter(category == "sex offense") |> 
  group_by(category, month, population) |> 
  summarise(pg_total = n()) |> 
  mutate(pg_sexoffense_rate = pg_total/population * 100000)

moco_sexoffense <- new_mococrime |> 
  filter(category == "sex offense") |> 
  group_by(category, month, population) |> 
  summarise(moco_total = n()) |> 
  mutate(moco_sexoffense_rate = moco_total/population * 100000)

combined_sexoffense <- pg_sexoffense |> 
  inner_join(moco_sexoffense, join_by(category, month))



pg_robbery <- new_pgcrime |> 
  filter(category == "robbery") |> 
  group_by(category, month, population) |> 
  summarise(pg_total = n()) |> 
  mutate(pg_robbery_rate = pg_total/population * 100000)

moco_robbery <- new_mococrime |> 
  filter(category == "robbery") |> 
  group_by(category, month, population) |> 
  summarise(moco_total = n()) |> 
  mutate(moco_robbery_rate = moco_total/population * 100000)

combined_robbery <- pg_robbery |> inner_join(moco_robbery, join_by(category, month))



pg_burglary <- new_pgcrime |> 
  filter(category == "burglary") |> 
  group_by(category, month, population) |> 
  summarise(pg_total = n()) |> 
  mutate(pg_burglary_rate = pg_total/population * 100000)

moco_burglary <- new_mococrime |> 
  filter(category == "burglary") |> 
  group_by(category, month, population) |> 
  summarise(moco_total = n()) |> 
  mutate(moco_burglary_rate = moco_total/population * 100000)

combined_burglary <- pg_burglary |> 
  inner_join(moco_burglary, join_by(category, month))



pg_autotheft <- new_pgcrime |> 
  filter(category == "auto theft") |> 
  group_by(category, month, population) |> 
  summarise(pg_total = n()) |> 
  mutate(pg_autotheft_rate = pg_total/population * 100000)

moco_autotheft <- new_mococrime |> 
  filter(category == "auto theft") |> 
  group_by(category, month, population) |> 
  summarise(moco_total = n()) |> 
  mutate(moco_autotheft_rate = moco_total/population * 100000)

combined_autotheft <- pg_autotheft |> 
  inner_join(moco_autotheft, join_by(category, month))



pg_theftfromauto <- new_pgcrime |> 
  filter(category == "theft from auto") |> 
  group_by(category, month, population) |> 
  summarise(pg_total = n()) |> 
  mutate(pg_theftfromauto_rate = pg_total/population * 100000)

moco_theftfromauto <- new_mococrime |> 
  filter(category == "theft from auto") |> 
  group_by(category, month, population) |> 
  summarise(moco_total = n()) |> 
  mutate(moco_theftfromauto_rate = moco_total/population * 100000)

combined_theftfromauto <- pg_theftfromauto |> 
  inner_join(moco_theftfromauto, join_by(category, month))



pg_homicide <- new_pgcrime |> 
  filter(category == "homicide") |> 
  group_by(category, month, population) |> 
  summarise(pg_total = n()) |> 
  mutate(pg_homicide_rate = pg_total/population * 100000)

moco_homicide <- new_mococrime |> 
  filter(category == "homicide") |> 
  group_by(category, month, population) |> 
  summarise(moco_total = n()) |> 
  mutate(moco_homicide_rate = moco_total/population * 100000)

combined_homicide <- pg_homicide |> 
  inner_join(moco_homicide, join_by(category, month))



pg_theft <- new_pgcrime |> 
  filter(category == "theft") |> 
  group_by(category, month, population) |> 
  summarise(pg_total = n()) |> 
  mutate(pg_theft_rate = pg_total/population * 100000)

moco_theft <- new_mococrime |> 
  filter(category == "theft") |> 
  group_by(category, month, population) |> 
  summarise(moco_total = n()) |> 
  mutate(moco_theft_rate = moco_total/population * 100000)

combined_theft <- pg_theft |> 
  inner_join(moco_theft, join_by(category, month))
```



## I compared the rates for each crime between the two counties in a graph so I could more easily analyze the data. Additionally, I color-coded the data to make it easier to differentiate and added titles, labels and captions.

```{r}
combined_assault |> 
  ggplot() +
  geom_line((aes(x=month, y=pg_assault_rate)), color = "red", linewidth = 1) +
  geom_line(aes(x=month, y=moco_assault_rate), color = "blue", linewidth = 1) +
  labs(title = "Assault rates in Montgomery County, Maryland, consistently higher than in neighboring Prince George's County",
       x = "Month",
       y = "Assault rate per 10,000 residents",
       caption = "Sources: Montgomery and Prince George's counties' crime archives")


combined_autotheft |> 
  ggplot() +
  geom_line((aes(x=month, y=pg_autotheft_rate)), color = "red", linewidth = 1) +
  geom_line(aes(x=month, y=moco_autotheft_rate), color = "blue", linewidth = 1) +
  labs(title = "Auto theft rates in Prince George's County, Maryland, consistently higher than in neighboring Montgomery County",
       x = "Month",
       y = "Auto theft rate per 10,000 residents",
       caption = "Sources: Montgomery and Prince George's counties' crime archives")


combined_burglary |> 
  ggplot() +
  geom_line((aes(x=month, y=pg_burglary_rate)), color = "red", linewidth = 1) +
  geom_line(aes(x=month, y=moco_burglary_rate), color = "blue", linewidth = 1) +
  labs(title = "Burglary rates in Prince George's and Montgomery counties in Maryland have decreased since 2017",
       x = "Month",
       y = "Burglary rate per 10,000 residents",
       caption = "Sources: Montgomery and Prince George's counties' crime archives")


combined_homicide |> 
  ggplot() +
  geom_line((aes(x=month, y=pg_homicide_rate)), color = "red", linewidth = 1) +
  geom_line(aes(x=month, y=moco_homicide_rate), color = "blue", linewidth = 1) +
  labs(title = "Homicide rates in Prince George's County, Maryland, consistently higher than in neighboring Montgomery County",
       x = "Month",
       y = "Homicide rate per 10,000 residents",
       caption = "Sources: Montgomery and Prince George's counties' crime archives")


combined_robbery |> 
  ggplot() +
  geom_line((aes(x=month, y=pg_robbery_rate)), color = "red", linewidth = 1) +
  geom_line(aes(x=month, y=moco_robbery_rate), color = "blue", linewidth = 1) +
  labs(title = "Robbery rates in Prince George's County, Maryland, consistently higher than in neighboring Montgomery County",
       x = "Month",
       y = "Robbery rate per 10,000 residents",
       caption = "Sources: Montgomery and Prince George's counties' crime archives")


combined_sexoffense |> 
  ggplot() +
  geom_line((aes(x=month, y=pg_sexoffense_rate)), color = "red", linewidth = 1) +
  geom_line(aes(x=month, y=moco_sexoffense_rate), color = "blue", linewidth = 1) +
  labs(title = "Sex offense rates in Montgomery County, Maryland, generally higher than in neighboring Prince George's County",
       x = "Month",
       y = "Sex offense rate per 10,000 residents",
       caption = "Sources: Montgomery and Prince George's counties' crime archives")


combined_theft |> 
  ggplot() +
  geom_line((aes(x=month, y=pg_theft_rate)), color = "red", linewidth = 1) +
  geom_line(aes(x=month, y=moco_theft_rate), color = "blue", linewidth = 1) +
  labs(title = "Theft rates in Montgomery County, Maryland, generally higher than in neighboring Prince George's County",
       x = "Month",
       y = "Theft rate per 10,000 residents",
       caption = "Sources: Montgomery and Prince George's counties' crime archives")


combined_theftfromauto |> 
  ggplot() +
  geom_line((aes(x=month, y=pg_theftfromauto_rate)), color = "red", linewidth = 1) +
  geom_line(aes(x=month, y=moco_theftfromauto_rate), color = "blue", linewidth = 1) +
  labs(title = "Theft from automobile rates in Prince George's and Montgomery counties in Maryland have decreased since 2017",
       x = "Month",
       y = "Theft from automobile rate per 10,000 residents",
       caption = "Sources: Montgomery and Prince George's counties' crime archives")
```



## Calculating unusual z-scores


1. Calculate the correlations between assault and sex offenses, etc.

2. Calculate the monthly means for each crime and analyze for any unusual z-scores.

3. Analyze any increases or decreases in crime rates and evaluate for statistical significance.




Hypotheses:

1. Montgomery County has higher rates of burglaries, automobile thefts and thefts from automobiles than Prince George's County

2. Prince George's County has higher rates of assault, sex offenses and overall theft than Montgomery County

3. Prince George's County has higher overall rates of these six crime categories than Montgomery County


Possible tests:

1. Calculating correlations among the different crime categories I am analyzing.

2. Calculating the correlation between assault and overall crime rates for Prince George's and Montgomery counties

3. Calculating the z-scores of the number of crimes committed per month and analyzing any unusual z-scores.

4. Calculating whether any decreases or increases in crime were statistically significant. If I found that they were, I could also calculate their effect size. 
