---
title: "project_test"
output: html_document
---

```{r}
library(tidyverse)

knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(knitr) 

library(lubridate)

library(dplyr)

library(scales)
```




```{r}
pgcrime17_23 <- read_csv("PGCRIME17-23.csv")

pgcrime23_26 <- read_csv("PGCRIME23-26.csv")

mococrime <- read_csv("MOCOCRIME.zip")

pgcrime <- bind_rows(pgcrime17_23, pgcrime23_26)
```



## I sorted all the crimes into seven categories (assault, burglary, auto theft, theft from auto, robbery, sex offenses and homicide) to facilitate my analysis.

```{r}
pgcrime <- pgcrime |> 
  mutate(category = case_when(
    str_detect(`Clearance Code Inc Type`, "ASSAULT") ~ "assault",
    str_detect(`Clearance Code Inc Type`, "ASSAULT, WEAPON") ~ "assault",
    str_detect(`Clearance Code Inc Type`, "ASSAULT, SHOOTING") ~ "assault",
    str_detect(`Clearance Code Inc Type`, "B & E, RESIDENTIAL") ~ "burglary",
    str_detect(`Clearance Code Inc Type`, "B & E, RESIDENTIAL (VACANT)") ~ "burglary",
    str_detect(`Clearance Code Inc Type`, "B & E, COMMERCIAL") ~ "burglary",
    str_detect(`Clearance Code Inc Type`, "B & E, OTHER") ~ "burglary",
    str_detect(`Clearance Code Inc Type`, "AUTO, STOLEN") ~ "auto theft",
    str_detect(`Clearance Code Inc Type`, "AUTO, STOLEN & RECOVERED") ~ "auto theft",
    str_detect(`Clearance Code Inc Type`, "ROBBERY, OTHER") ~ "robbery",
    str_detect(`Clearance Code Inc Type`, "ROBBERY, COMMERCIAL") ~ "robbery",
    str_detect(`Clearance Code Inc Type`, "ROBBERY, RESIDENTIAL") ~ "robbery",
    str_detect(`Clearance Code Inc Type`, "ROBBERY, VEHICLE") ~ "robbery",
    str_detect(`Clearance Code Inc Type`, "ROBBERY, VEHICLE") ~ "robbery",
    str_detect(`Clearance Code Inc Type`, "THEFT FROM AUTO") ~ "theft from auto",
    str_detect(`Clearance Code Inc Type`, "THEFT") ~ "theft",
    str_detect(`Clearance Code Inc Type`, "SEX OFFENSE") ~ "sex offense",
    str_detect(`Clearance Code Inc Type`, "HOMICIDE") ~ "homicide",
    .default = NA
  )
)


mococrime <- mococrime |>
  mutate(category = case_when(
    str_detect(`Crime Name3`, "ASSAULT - 2ND DEGREE") ~ "assault",
    str_detect(`Crime Name3`, "ASSAULT - INTIMIDATION (INCLUDES STALKING)") ~ "assault",
    str_detect(`Crime Name3`, "ASSAULT - AGGRAVATED - PUB OFF-OTHER WEAPON") ~ "assault",
    str_detect(`Crime Name3`, "ASSAULT - AGGRAVATED - POL OFF-OTHER WEAPON") ~ "assault",
    str_detect(`Crime Name3`, "ASSAULT - AGGRAVATED - FAMILY-GUN") ~ "assault",
    str_detect(`Crime Name3`, "ASSAULT - SIMPLE") ~ "assault",
    str_detect(`Crime Name3`, "ASSAULT - AGGRAVATED - GUN") ~ "assault",
    str_detect(`Crime Name3`, "ASSAULT - AGGRAVATED - NON-FAMILY-OTHER WEAPON") ~ "assault",
    str_detect(`Crime Name3`, "ASSAULT - AGGRAVATED - FAMILY-STRONG-ARM") ~ "assault",
    str_detect(`Crime Name3`, "ASSAULT - AGGRAVATED - NON-FAMILY-GUN") ~ "assault",
    str_detect(`Crime Name3`, "ASSAULT - AGGRAVATED - NON-FAMILY-STRONG-ARM") ~ "assault",
    str_detect(`Crime Name3`, "ATTEMPTED ASSAULT AGGRAVATED POL OFF KNIFE") ~ "assault",
    str_detect(`Crime Name3`, "ASSAULT - AGGRAVATED - POL OFF-STRONG-ARM") ~ "assault",
    str_detect(`Crime Name3`, "ASSAULT - AGGRAVATED - POL OFF-GUN") ~ "assault",
    str_detect(`Crime Name3`, "ASSAULT - AGGRAVATED - PUB OFF-STRONG-ARM") ~ "assault",
    str_detect(`Crime Name3`, "ASSAULT - AGGRAVATED - PUB OFF-GUN") ~ "assault",
    
    .default = NA
  )
)

mococrime <- mococrime |>
  mutate(category = case_when(
    str_detect(`Crime Name3`, "SEX OFFENSE - AGAINST CHILD-FONDLING") ~ "sex offense",
    str_detect(`Crime Name3`, "SEX ASSAULT - FONDLING") ~ "sex offense",
    str_detect(`Crime Name3`, "SEX OFFENSE - PEEPING TOM") ~ "sex offense",
    str_detect(`Crime Name3`, "SEX OFFENSE - INDECENT EXPOSURE") ~ "sex offense",
    str_detect(`Crime Name3`, "COMM SEX OFF - KEEPING HOUSE ILL FAME") ~ "sex offense",
    str_detect(`Crime Name3`, "SEX OFFENSE - INCEST WITH ADULT") ~ "sex offense",
    str_detect(`Crime Name3`, "SEX ASSAULT - SODOMY-GIRL-GUN") ~ "sex offense",
    str_detect(`Crime Name3`, "SEX ASSAULT - SODOMY-GIRL-WEAPON") ~ "sex offense",
    str_detect(`Crime Name3`, "COMM SEX OFF - PROCURE FOR PROSTITUTE (PIMPING)") ~ "sex offense",
    str_detect(`Crime Name3`, "COMM SEX OFF - PROCURE PROSTITUTE - MINOR") ~ "sex offense",
    str_detect(`Crime Name3`, "SEX ASSAULT - SODOMY-WOMAN-GUN") ~ "sex offense",
    str_detect(`Crime Name3`, "COMM SEX OFF - PROSTITUTION") ~ "sex offense",
    str_detect(`Crime Name3`, "HOMOSEXUAL ACT WITH BOY") ~ "sex offense",
    str_detect(`Crime Name3`, "SEX OFFENSE - INCEST WITH MINOR") ~ "sex offense",
    str_detect(`Crime Name3`, "SEX ASSAULT - SODOMY-MAN-WEAPON") ~ "sex offense",
    str_detect(`Crime Name3`, "COMM SEX OFF - HOMOSEXUAL PROSTITUTION") ~ "sex offense",
    str_detect(`Crime Name3`, "SEX ASSAULT - SODOMY-WOMAN-WEAPON") ~ "sex offense",
    str_detect(`Crime Name3`, "SEX ASSAULT - SODOMY-MAN-GUN") ~ "sex offense",
    str_detect(`Crime Name3`, "COMM SEX OFF - TRANSPORT FEMALE INTERSTATE FOR IMM") ~ "sex offense",
    .default = category
  )
)

mococrime <- mococrime |>
  mutate(category = case_when(
    str_detect(`Crime Name3`, "ROBBERY - STREET-GUN") ~ "robbery",
    str_detect(`Crime Name3`, "ROBBERY - RESIDENTIAL-STRONG-ARM") ~ "robbery",
    str_detect(`Crime Name3`, "ROBBERY - STREET-STRONG-ARM") ~ "robbery",
    str_detect(`Crime Name3`, "ROBBERY - STREET-OTHER WEAPON") ~ "robbery",
    str_detect(`Crime Name3`, "ROBBERY - BUSINESS-GUN") ~ "robbery",
    str_detect(`Crime Name3`, "ROBBERY - FORCIBLE PURSE SNATCHING") ~ "robbery",
    str_detect(`Crime Name3`, "ROBBERY (DESCRIBE OFFENSE)") ~ "robbery",
    str_detect(`Crime Name3`, "ROBBERY - RESIDENTIAL-GUN") ~ "robbery",
    str_detect(`Crime Name3`, "ROBBERY - BUSINESS-OTHER WEAPON") ~ "robbery",
    str_detect(`Crime Name3`, "ROBBERY - CARJACKING - ARMED") ~ "robbery",
    str_detect(`Crime Name3`, "ROBBERY - RESIDENTIAL-OTHER WEAPON") ~ "robbery",
    str_detect(`Crime Name3`, "ROBBERY - BANKING-TYPE INST") ~ "robbery",
    .default = category
  )
)

mococrime <- mococrime |>
  mutate(category = case_when(
    str_detect(`Crime Name3`, "BURGLARY - SAFE-VAULT") ~ "burglary",
    str_detect(`Crime Name3`, "BURGLARY - BANKING-TYPE INST") ~ "burglary",
    str_detect(`Crime Name3`, "BURGLARY (DESCRIBE OFFENSE)") ~ "burglary",
    str_detect(`Crime Name3`, "BURGLARY - NO FORCED ENTRY-NONRESIDENTIAL") ~ "burglary",
    str_detect(`Crime Name3`, "BURGLARY - FORCED ENTRY-RESIDENTIAL") ~ "burglary",
    str_detect(`Crime Name3`, "BURGLARY - FORCED ENTRY-NONRESIDENTIAL") ~ "burglary",
    str_detect(`Crime Name3`, "BURGLARY - NO FORCED ENTRY-RESIDENTIAL") ~ "burglary",
    .default = category
  )
)

mococrime <- mococrime |>
  mutate(category = case_when(
    str_detect(`Crime Name3`, "STOLEN VEHICLE (DESCRIBE OFFENSE)") ~ "auto theft",
    str_detect(`Crime Name3`, "AUTO THEFT - VEHICLE THEFT") ~ "auto theft",
    str_detect(`Crime Name3`, "AUTO THEFT - THEFT VEHICLE BY BAILEE") ~ "auto theft",
    str_detect(`Crime Name3`, "AUTO THEFT - STRIP STOLEN VEHICLE") ~ "auto theft",
    str_detect(`Crime Name3`, "AUTO THEFT - THEFT AND STRIP VEHICLE") ~ "auto theft",
    str_detect(`Crime Name3`, "AUTO THEFT - THEFT AND SALE VEHICLE") ~ "auto theft",
    str_detect(`Crime Name3`, "AUTO THEFT - THEFT AND USE VEHICLE OTHER CRIME") ~ "auto theft",
    .default = category
  )
)

mococrime <- mococrime |>
  mutate(category = case_when(
    str_detect(`Crime Name3`, "LARCENY - FROM AUTO") ~ "theft from auto",
    str_detect(`Crime Name3`, "LARCENY - AUTO PARTS") ~ "theft from auto",
    .default = category
  )
)

mococrime <- mococrime |>
  mutate(category = case_when(
    str_detect(`Crime Name3`, "HOMICIDE - WILLFUL KILL-GUN") ~ "homicide",
    str_detect(`Crime Name3`, "HOMICIDE (DESCRIBE OFFENSE)") ~ "homicide",
    str_detect(`Crime Name3`, "HOMICIDE - WILLFUL KILL") ~ "homicide",
    str_detect(`Crime Name3`, "HOMICIDE - WILLFUL KILL-FAMILY") ~ "homicide",
    str_detect(`Crime Name3`, "HOMICIDE - JUSTIFIABLE") ~ "homicide",
    str_detect(`Crime Name3`, "HOMICIDE - WILLFUL KILL-FAMILY-GUN") ~ "homicide",
    str_detect(`Crime Name3`, "HOMICIDE - WILLFUL KILL-NON-FAMILY-GUN") ~ "homicide",
    str_detect(`Crime Name3`, "HOMICIDE - WILLFUL KILL-NON-FAMILY") ~ "homicide",
    str_detect(`Crime Name3`, "HOMICIDE - NEGLIGENT MANSLAUGHTER") ~ "homicide",
    str_detect(`Crime Name3`, "HOMICIDE - BY NON-POLICE - JUSTIFIABLE") ~ "homicide",
    str_detect(`Crime Name3`, "HOMICIDE - JOHN OR JANE DOE - NO WARRANT") ~ "homicide",
    .default = category
  )
)

mococrime <- mococrime |>
  mutate(category = case_when(
    str_detect(`Crime Name3`, "LARCENY (DESCRIBE OFFENSE)") ~ "theft",
    str_detect(`Crime Name3`, "LARCENY - POCKET PICKING") ~ "theft",
    str_detect(`Crime Name3`, "LARCENY - FROM INTERSTATE SHIPMENT") ~ "theft",
    str_detect(`Crime Name3`, "LARCENY - THEFT OF US GOVERNMENT PROPERTY") ~ "theft",
    str_detect(`Crime Name3`, "LARCENY - FROM COIN MACHINE") ~ "theft",
    str_detect(`Crime Name3`, "LARCENY - FROM SHIPMENT") ~ "theft",
    str_detect(`Crime Name3`, "LARCENY - FROM BANKING-TYPE INST") ~ "theft",
    str_detect(`Crime Name3`, "LARCENY - FROM MALLS") ~ "theft",
    str_detect(`Crime Name3`, "LARCENY - FROM YARDS") ~ "theft",
    str_detect(`Crime Name3`, "LARCENY - PURSE SNATCHING - NO FORCE") ~ "theft",
    str_detect(`Crime Name3`, "LARCENY - POCKET PICKING") ~ "theft",
    str_detect(`Crime Name3`, "LARCENY - POSTAL") ~ "theft",
    str_detect(`Crime Name3`, "LARCENY - FROM BLDG") ~ "theft",
    str_detect(`Crime Name3`, "LARCENY - SHOPLIFTING") ~ "theft",
    .default = category
  )
)
```





```{r}
pgcrime <- pgcrime |>
  mutate(Date = mdy(Date)) |> 
  mutate(
    month = floor_date(Date, "month")
  )
```

```{r}
mococrime <- mococrime |> 
  mutate(Date = date(mdy_hm(Start_Date_Time)))|>
      mutate(month = floor_date(Date, "month"))
```




```{r}
new_pgcrime <- pgcrime |> filter(Date < '2026-04-01', Date > '2017-03-31')
```

```{r}
new_mococrime <- mococrime |> filter(Date < '2026-04-01', Date > '2017-03-31')
```



```{r}
new_mococrime <- new_mococrime %>%
  mutate(population = 1053067)

new_pgcrime <- new_pgcrime %>%
  mutate(population = 946971)
```


```{r}
pg_assault <- new_pgcrime |> 
  filter(category == "assault") |> 
  group_by(category, month, population) |> 
  summarise(pg_total = n()) |> 
  mutate(pg_assault_rate = pg_total/population * 100000)

moco_assault <- new_mococrime |> 
  filter(category == "assault") |> 
  group_by(category, month, population) |> 
  summarise(moco_total = n()) |> 
  mutate(moco_assault_rate = moco_total/population * 100000)

combined_assault <- pg_assault |> 
  inner_join(moco_assault, join_by(category, month))



pg_sexoffense <- new_pgcrime |> 
  filter(category == "sex offense") |> 
  group_by(category, month, population) |> 
  summarise(pg_total = n()) |> 
  mutate(pg_sexoffense_rate = pg_total/population * 100000)

moco_sexoffense <- new_mococrime |> 
  filter(category == "sex offense") |> 
  group_by(category, month, population) |> 
  summarise(moco_total = n()) |> 
  mutate(moco_sexoffense_rate = moco_total/population * 100000)

combined_sexoffense <- pg_sexoffense |> 
  inner_join(moco_sexoffense, join_by(category, month))



pg_robbery <- new_pgcrime |> 
  filter(category == "robbery") |> 
  group_by(category, month, population) |> 
  summarise(pg_total = n()) |> 
  mutate(pg_robbery_rate = pg_total/population * 100000)

moco_robbery <- new_mococrime |> 
  filter(category == "robbery") |> 
  group_by(category, month, population) |> 
  summarise(moco_total = n()) |> 
  mutate(moco_robbery_rate = moco_total/population * 100000)

combined_robbery <- pg_robbery |> inner_join(moco_robbery, join_by(category, month))



pg_burglary <- new_pgcrime |> 
  filter(category == "burglary") |> 
  group_by(category, month, population) |> 
  summarise(pg_total = n()) |> 
  mutate(pg_burglary_rate = pg_total/population * 100000)

moco_burglary <- new_mococrime |> 
  filter(category == "burglary") |> 
  group_by(category, month, population) |> 
  summarise(moco_total = n()) |> 
  mutate(moco_burglary_rate = moco_total/population * 100000)

combined_burglary <- pg_burglary |> 
  inner_join(moco_burglary, join_by(category, month))



pg_autotheft <- new_pgcrime |> 
  filter(category == "auto theft") |> 
  group_by(category, month, population) |> 
  summarise(pg_total = n()) |> 
  mutate(pg_autotheft_rate = pg_total/population * 100000)

moco_autotheft <- new_mococrime |> 
  filter(category == "auto theft") |> 
  group_by(category, month, population) |> 
  summarise(moco_total = n()) |> 
  mutate(moco_autotheft_rate = moco_total/population * 100000)

combined_autotheft <- pg_autotheft |> 
  inner_join(moco_autotheft, join_by(category, month))



pg_theftfromauto <- new_pgcrime |> 
  filter(category == "theft from auto") |> 
  group_by(category, month, population) |> 
  summarise(pg_total = n()) |> 
  mutate(pg_theftfromauto_rate = pg_total/population * 100000)

moco_theftfromauto <- new_mococrime |> 
  filter(category == "theft from auto") |> 
  group_by(category, month, population) |> 
  summarise(moco_total = n()) |> 
  mutate(moco_theftfromauto_rate = moco_total/population * 100000)

combined_theftfromauto <- pg_theftfromauto |> 
  inner_join(moco_theftfromauto, join_by(category, month))



pg_homicide <- new_pgcrime |> 
  filter(category == "homicide") |> 
  group_by(category, month, population) |> 
  summarise(pg_total = n()) |> 
  mutate(pg_homicide_rate = pg_total/population * 100000)

moco_homicide <- new_mococrime |> 
  filter(category == "homicide") |> 
  group_by(category, month, population) |> 
  summarise(moco_total = n()) |> 
  mutate(moco_homicide_rate = moco_total/population * 100000)

combined_homicide <- pg_homicide |> 
  inner_join(moco_homicide, join_by(category, month))



pg_theft <- new_pgcrime |> 
  filter(category == "theft") |> 
  group_by(category, month, population) |> 
  summarise(pg_total = n()) |> 
  mutate(pg_theft_rate = pg_total/population * 100000)

moco_theft <- new_mococrime |> 
  filter(category == "theft") |> 
  group_by(category, month, population) |> 
  summarise(moco_total = n()) |> 
  mutate(moco_theft_rate = moco_total/population * 100000)

combined_theft <- pg_theft |> 
  inner_join(moco_theft, join_by(category, month))
```



## The result that surprised me the most from the graphic was Montgomery County's high assault rate compared to Prince George's County. It went against my hypothesis and is an outlier compared to the other graphs, where the crime rates are closer together. I suspected that my categorization of assault may have contributed to this discrepancy. 

## It appears that automobile theft rates in both counties spiked in 2023. I have been hearing more about automobile thefts in the past few years, so this piqued my interest and will warrant further research.

## It appears that burglary rates in both counties have decreased overall since April 2017. This decrease warrants further research.

## It appears that homicide rates in Prince George's County spiked in late 2020 and in 2023. These spikes warrant further research.

## It appears that robbery rates in Prince George's County spiked in late 2023 and have steadily decreased since then. This change warrants further research.

## It appears that Prince George's County had a relatively high sex offense rate in 2017. However, since that year, both counties have sustained sex offense rates of less than five per 10,000 residents. However, to determine whether this number is high or low, I will have to compare it to the national average.

## It appears that theft rates in both counties had a spike in late 2024. This spike warrants further research.

## Theft from automobile rates in both counties appear to zigzag without a consistent pattern. This lack of a pattern warrants further research.


## The spike in auto thefts in both counties that I noticed in the graph was corroborated by my z-score analysis, further proving that said spike was unusual and underlining the need for additional research. 

## The anomalies in burglary rates in Prince George's County caught my eye because the graphs appeared to show an overall decrease. I will need to research this further to detect the causes of these anomalies.

## The spike in homicide rates in Prince George's County that I noticed in the graph are corroborated by my z-score analysis, indicating a need for further research. 

## The spike in robbery rates in Prince George's County that I noticed in the graph was corroborated by my z-score analysis, indicating a need for further research. 

## The spike in sex offense rates in Prince George's County that I noticed in the graph was corroborated by my z-score analysis, indicating a need for further research. 

## The spike in theft rates in both counties that I noticed in the graph was not corroborated by my z-score analysis. However, those spikes may still be signficant, and I still need to research them.


## The result that fascinated me the most was the high correlation between automobile theft rates in both counties. Though I had noticed that the rates between the counties almost mirrored each other when observing the graph, I forgot to mention it. However, when thought about in the context of the real world, it makes sense. Neighboring counties, to an extent, should experience similar increases and decreases in crime rates. Based on this viewpoint, the lack of correlation between any other crime category is more surprising.
